### 2024-08-29 assignments/research_proposal/response_to_reviewers.Rmd
add lectures 1-5

---
title: |
        | Assignment: Response to Reviewers
subtitle: |
        | Ph.D. Environmental and Development Economics
        | UMN Applied Economics 2024
        | Instructor: Raahil Madhok
# Output type and options (no TOC and yes fig captions)
output:
  bookdown::pdf_document2: 
    toc: false
    fig_caption: yes
    # This also keeps the .tex file that is used to make the output (we will eventually need this for publication)
    keep_tex: yes

# This includes latex arguments
header-includes:
  - \input{header.tex}

---

**Due Date: October 10th, 2024**

# Overview
In this assignment, you will write a mock ``response to the reviewer'' based on the peer review you received for your research proposal. This process mimics the response letter you would write in a journal submission, where you address each comment made by the referee, explain how you revised the paper, and justify any decisions you made in response to the feedback.

The purpose of this exercise is to help you practice responding to critical feedback in a constructive and professional manner, while also improving your proposal based on your peer's suggestions.

# Structure of the Response

Your response should include the following components:

## Introduction Paragraph

- **Thank the reviewer**: Begin by thanking your reviewer for their thoughtful feedback
- **General summary**: Summarize the key points of the review and provide a brief overview of the changes you have made to your proposal as a result of the feedback.
- **Tone**: Keep your tone respectful, professional, and appreciative.

## Point-by-Point Response
For each comment made by the reviewer, provide a clear and detailed response. The unwritten rule is to copy and paste the peer review into a document, turn each comment into a bulletted list, and then respond underneath each comment in the list.

For each point:

- **Restate the reviewer’s comment**: Begin by restating the comment or suggestion. This ensures clarity and demonstrates that you fully understand the critique.

- **Explain your response**:
    - **If you have made changes**: Explain what specific changes you made to the proposal in response to the reviewer’s comment. Clearly identify where in the revised proposal these changes can be found (e.g., “We have revised Section 2, paragraphs 3 and 4, to clarify the theoretical framework and the role of environmental externalities.”).
    - **If you disagree with the comment**: Provide a reasoned explanation for why you have chosen not to make the suggested change. Be polite and respectful, offering evidence or justification for maintaining your original approach. For example, “While we appreciate this suggestion, we believe that keeping the focus on District-level data is more consistent with the existing literature.”

- **Detail the improvement**: Explain how the changes you made improve the overall quality and clarity of your proposal. This will help the reviewer see the value of the revisions.

## Conclusion Paragraph

- **Reaffirm gratitude**: Reiterate your appreciation for the reviewer’s input.
- **Summary of revisions**: Offer a final summary of the revisions made and express your confidence that the proposal has been strengthened as a result.

# Guidelines for Writing Your Response

- **Be thorough**: Ensure that every comment in the review, whether major or minor, receives a response. Do not skip any points.

- **Be specific**: When describing the changes you made, be as specific as possible. Reference the exact sections or pages of your proposal where you incorporated changes.

- **Be polite and constructive**: Even if you disagree with some of the reviewer’s suggestions, maintain a respectful tone. Use evidence and logical reasoning to support your decisions.

- **Reflect improvements**: Make sure the changes you discuss in your response are actually incorporated into the revised version of your proposal. Your response and your revised proposal should be fully aligned.

# Format

- **Length:** Responses (not including the original comment) should not exceed 4 pages (1000 words)

- **Format:** 12 point font, double spaced, 1 inch margins

- **Submission:** PDF document (latex or word)

# Rubric
\begin{table}[h!]
\centering
\begin{tabular}{|l|p{8cm}|c|c|}
\hline
\textbf{Category} & \textbf{Criteria} & \textbf{Points} & \textbf{Score} \\
\hline
\multirow{2}{*}{\textbf{Comprehensive Response}} & Addresses all major and minor points raised by the reviewer, with no comments left unaddressed. & 10 & \\
\hline
\multirow{2}{*}{\textbf{Clarity and Specificity}} & Clearly provides a detailed, specific response to each comment. & 10 & \\
\hline
\multirow{2}{*}{\textbf{Incorporation of Changes}} & Demonstrates thoughtful incorporation of reviewer suggestions into the proposal, with clear references to specific changes made. & 30 & \\
\hline
\multirow{2}{*}{\textbf{Justification for Disagreements}} & Provides clear and well-reasoned explanations for why certain suggestions were not incorporated, using evidence and logical reasoning. & 30 & \\
\hline
\multirow{2}{*}{\textbf{Tone and Professionalism}} & Maintains a respectful, professional, and constructive tone throughout the response, even when disagreeing. & 10 & \\
\hline
\multirow{2}{*}{\textbf{Writing Quality}} & The response document is well-organized, with clear writing, correct grammar, and appropriate length. & 10 & \\
\hline
\multicolumn{3}{|r|}{\textbf{Total Points Possible:}} & \textbf{100} \\
\hline
\end{tabular}
\end{table}