### 2024-09-02 lectures/module4/module4.Rmd
before first lecture

---
title: "| Environmental and Development Economics\n| Module 4 - WTP for Environmental Quality
  \n"
author: |
  | Raahil Madhok
  | UMN Apllied Economics
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Montpellier
  slidy_presentation: default
  ioslides_presentation: default
classoption: aspectratio=169
header-includes:
- "\\AtBeginDocument{\\title[Module 4 - WTP for environment]{Environmental and Development
  Economics \\\\ Module 4 - WTP for Environmental Quality in LMICs}}"
- \input{header.tex}
- \usepackage{gensymb}
---

## 
\centering
\Huge Lecture 6

\large WTP for Environmental Quality in Developing Countries

## Housekeeping

- State things that are due.

## Today

- **Guiding question:** what is the WTP for environmental quality in LMICs?
\vfill
- Today's focus: How do we even measure this?
  - Behavioral models
  - Revealed preference approaches

\vfill
- Next time: why is WTP so low in developing countries?

## Measuring Willingness To Pay

- Slides on Logit and Mixed Logit

## 
\huge How is this implemented in practice?

## Kremer et al. (2011): WTP for Clean Water in Kenya

- **Research Question:** What is household WTP for clean water in LICs?
  - What are health impacts of water quality improvements? (we will skip this)
\vfill
- **Setting:** 1,354 households in Kenya
\vfill
- **Intervention:** Improve water quality at source for 184 natural springs
\vfill
- **Empirical Strategy:** RCT + Revealed preference valuation
\vfill
- **Results:** Clean water valued at \$USD 2.96 per hh (\$800 VSL)

## Spring Protection

- Spring water is crucial drinking water source in Africa
  - But vulnerable to contamination when people dip vessels
  - Runoff introduces animal/human waste
\vfill
- Spring protection common in Africa to improve water quality at source
\vfill
- Seal off spring at source by encasing in concrete
  - Water flows out of pipe rather than ground, avoiding contamination
\vfill
- Nothing can ``break'' $\rightarrow$ less maintenance than wells/pumps 
\vfill
- How much are hh WTP for spring protection?

## Spring Protection
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "68%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/spring_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Spring Protection
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "75%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/spring_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Spring Protection
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/spring_fig3.png",
                        auto_pdf = TRUE
                        )

```

## Spring Protection
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "65%", fig.align = "center", fig.cap="Source: USAID"}

knitr::include_graphics(
                        path = "assets/spring_fig4.png",
                        auto_pdf = TRUE
                        )

```

## Study Sample
- Sample frame: list of all unprotected springs in Busia and Butere-Mumias districts
\vfill
- Treatment: 184 randomly selected for protection
  - 1,345 households selected that use springs
\vfill
- Phased design: all springs eventually protected
  - Round I (Jan-Apr 2005); 47 springs, 350 households
  - Round II (Apr-Nov 2005); 46 springs; 349 households
  - Round III (2005-2006); 91 springs; 685 households
\vfill
- 7-8 households from each spring select from ``user list''

## Data Collection

- Water quality measured at each spring based on EPA protocols
  - Measure: E.Coli (log of \# bacteria per 100ml)
\vfill
- Household survey
  - Diarrhea
  - Anthopometrics
  - Water collection behvavior
  - Socioeconomic data
\vfill
- No difference in baseline water quality between treatment/control springs

## Water Quality and Health
- Impact of spring protection on water quality:
$$W_{jt'}^{SP} = \alpha_{t} + \phi_{1}T_{jt} + X_{j}^{SP'}\phi_{2} + (T_{jt} \cdot X_{j}^{SP})'\phi_{3} + \epsilon_{jt}$$
  - where $W_{jt}$ is water quality for spring $j$ at time $t\in {0,1,2,3}$
  - $T_{jt}$ is treatment status (=1 after protection)
\vfill
- Impact on health
$$Y_{ijt} = \alpha_{i} + \alpha_{t} + \phi_{1} T_{jt} + X_{ij}'\phi_{2} + (T_{jt} \cdot X_{ij})'\phi_{3} + u_{ij} + \epsilon_{ijt}$$

  -  where $Y_{ijt}$ is whether child $i$ had diarrhea in the past week

## Results: Impact of Spring Protection on Water Quality and Health

**Water Quality**

- Huge improvement in water quality: 66\% $\downarrow$ in E.coli
- No heterogeneity by baseline hygiene knowledge or education
  - Hypothesis: educated hh may maintain spring better
- Improvement in home water quality: 24\% $\downarrow$ in E.coli 
  - Why is the effect smaller?

\vfill
**Health**

- Diarrhea incidence $\downarrow$ by 4.7pp (20\% reduction compared to control)
- No impact on child weight
- Positive effect on BMI
\vfill

**What if control households choose to use treatment springs?**

## Revealed Preference: Valuing Clean Water in Kenya

- Let valuation of water from source $j$ be $Z_{j}$
  - reflects health and non-health attributes (e.g. ease of collection)
\vfill
- $\beta_{i}$: indirect utility from protection ($T_{jt}$) at $j$ at time $t$
\vfill
- $C_{i}>0$ is time cost of household $i$
\vfill
- $C_{i}D_{ij}$: Travel cost for additional trip to $j$ where $D_{i}$ is round-trip distance
\vfill
- Each trip affected by unobserved factors (weather, queue, errands, mood)

## Revealed Preference Approach
- Household indirect utility from one trip to $j$ at time $t$ is:
$$ u_{ijt} = \beta_{i}T_{jt}+Z_{j}-C_{i}D_{ij} + e_{ijt}$$
  - where $e_{ijt}$ is iid type I extreme value distributed
\vfill
- $i$ chooses $j$ over $k$ if: 
$$\beta_{i}(T_{jt}-T_{kt})+(Z_{j}-Z_{k})-C_{i}(D_{ij}-D_{ik})+(e_{ijt}-e_{ikt}) \geq 0$$
- Using CDF of EV Type I error, prob. of choosing $j$ over alternatives $h$:
$$P(y_{ijt}|X) = \frac{exp(X_{ijt}^{'}\beta)}{\sum_{h}exp(X_{iht}^{'}\beta)} \equiv \rho_{ijt}$$

## Mixed Logit
- Follow-up survey gives list of alternative water sources
- What is ratio of coeff. on $T_{jt}$ to coeff. on walking time?
  <!-- - Value of spring protection in terms of minutes spent walking -->

- Add preference heterogeneity with mixed logit:
$$P(y_{ijt}|X) = \int_{\beta} \rho_{ijt}f(\beta)d\beta$$
  - where $f(\cdot)$ is the mixing distribution
- Maximize log-likelihood to estimate mean and s.d. of these distributions

## WTP results
- Conditional logit estimates: how do we interpret these?
  - Coefficient on walking distance ($C_{i}D_{ij}$) = -0.055
  - Coefficient on spring protection ($\beta_{i}$) = 0.51

- Can $\beta_{i}$ entangle other amenity values?
  <!-- - reduced need to bend down, lower collection times -->
  <!-- - yes, but when include perceived water quality, $\beta_{i}=0$, meaning the greater value placed on protected springs is due to impact on water quality, not other factors -->

- Households with kids under 3 find walking more costly
  - Coeff. on interaction b/w $T_{jt}$ and child is positive (insignificant)
  - Households with kids value water quality more

## WTP for spring protection

- Ratio of mixed logit coefficients: 2.95/0.21

- Households value clean water by willing to walk 14 extra minutes

- Average hh takes 48 trips per week (see Descriptive Stats)
- Total extra minutes per year = 14 x 48 x 52 = 35,078 minutes
- Assuming 8-hour workday $\rightarrow$ **mean value of clean water is 73 workdays**
  - Note: in paper, it is 32.4 workdays!

## Comparison with Stated Preference Values

## Converting WTP to dollars

**Survey evidence**

- Subset of 104 respondents
- Contingent valuation: are you WTP \$5 to walk 1 less minute? \$10?, etc 
- Divide WTP (in terms of time) by money value of time
- Regress time value on household characteristics, and predict on full sample
- Mean value of time = \$USD 0.088 per 8-hour day (7\% of ag wages)

\vfill

**Wages**

- \$USD 1.26 for 8-hour day (Suri, 2009)
- Time cost = 25\% of average workday wage

## Health Valuation and VSL

## Discussion
<!-- - what do you think about approach? -->
<!-- - what do you think about time valuation? -->
<!-- - Why not include income directly in indirect utility fxn? -->

## Distribution of Household Values of Clean Water (Mixed Logit)
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/kremer2011_fig1.png",
                        auto_pdf = TRUE
                        )

```
- Why is stated preference values larger? Why is CV distribution wider?