### 2024-11-11 lectures/module2/module2.Rmd
added .Rmd files

---
title: "| Environmental and Development Economics\n| Module 2 - How does development affect the environment?
  \n"
author: |
  | Raahil Madhok
  | UMN Applied Economics
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Montpellier
  slidy_presentation: default
  ioslides_presentation: default
classoption: aspectratio=169
header-includes:
- "\\AtBeginDocument{\\title[Module 2 - development and environment]{Environmental and Development
  Economics \\\\ Module 2 - How does development affect the environment?}}"
- \input{header.tex}
---

## 
\centering
\Huge Lecture 2

\large How does development affect the environment?

## Housekeeping

- Replication project (Ryan announcement)
  - Due Oct 17th
\vfill
- Research proposal
  - Start thinking about idea + come talk to me
  - First draft: Oct 3rd (pass/fail)
\vfill
- Lecture slides will be posted before each class
\vfill
- Reading list is a work in progress
\vfill
- 5 min break mid-way through lectures

## Today

- **Guiding question:** how does economic development affect the environment?
\vfill
- Descriptive overview
\vfill
- Channel I: Income effects (today)
  - **Changes in consumption**
  - **Energy**
  - Diet
\vfill
- Channel II: Technology and Infrastructure (next week)
  - Agricultural productivity
  - Infrastructure

\vfill
- Channel III: Institutions
  - Later in the course

\vfill

## How does economic development affect the environment?

- There is no one answer

- Choices that maximize economic growth often degrade environment

- But development expands our choice set (e.g. clean energy)

- Development also lowers $u'(c)$ (model from last week)

- Hence: the Environmental Kuznets Curve (EKC)

## Environmental Kuznets Curve

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "95%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ekc.png",
                        auto_pdf = TRUE
                        )

```

## Discussion: Do you believe EKC?

- Mechanisms?
\vfill

- Causality?
\vfill
- Robustness?
\vfill

## GDP and Lead in Gasoline
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ekc_lead.png",
                        auto_pdf = TRUE
                        )

```

## GDP and CO2
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "49%", fig.align = "center"}

knitr::include_graphics(
                        path = c("assets/ekc_co2.png","assets/ekc_legend.png"),
                        auto_pdf = TRUE
                        )

```

## GDP and PM2.5
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "49%", fig.align = "center"}

knitr::include_graphics(
                        path = c("assets/ekc_pm.png","assets/ekc_legend.png"),
                        auto_pdf = TRUE
                        )

```

## GDP and Ozone
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "49%", fig.align = "center"}

knitr::include_graphics(
                        path = c("assets/ekc_ozone.png","assets/ekc_legend.png"),
                        auto_pdf = TRUE
                        )

```

## My view

- Macro correlations give limited answer
\vfill
- Not the right question
  - What do we mean by development?
  - e.g. manufacturing $\uparrow$ pollution, but services may not

\vfill
- Narrow, micro/empirical papers more helpful
  - even if they don't find consensus

## Discussion: What is the ideal (quasi-) experiment?

- Can we randomize GDP?

- What variation in "development" can approximate the experiment?
  - what proxies development? (see \href{https://github.com/teamookla/ookla-open-data}{ookla speed test})
  - what does the proxy miss?
  - what (non-random) variation isolates causal relationship?

## Discussion: What about mechanisms?

- **Why and how** does economic development affect the environment?

- Theory is your friend: especially total derivatives...
  - Sometimes enough to rationalize mechanisms
  - Otherwise generate testable predictions

- Heterogeneity is your friend
  - But have a reason **why** you choose one channel over another

## Alix-Garcia et al. (2013): Ecological footprint of poverty alleviation

**Question:** What is the effect of income on environmental  quality?
\vfill

- Poverty alleviation may $\uparrow$ demand for resource-intensive good
\vfill
- Or, it may $\uparrow$ demand for conservation
\vfill
- **This paper:** Exploit RDD eligibility to study impact of CCT on forests

## Why can't we regress forests on income?

- Income correlated with other things that affect environment

\vfill
- Like what?
  <!-- population agricultural returns, forest product prices, capital -->

\vfill
- If we control for OVB, will panel fixed effects work? 
  <!-- - No: small year-to-year income changes do not reflect ``development'' -->
  <!-- - Predictable short-term income fluctuations different than permanent change -->

## Mexico's Oportunidades Program

- **Goal:** Ambitious program to build human capital

\vfill
- **Scope:** \$2.6 billion/year (half of anti-poverty budget)
  - Transfers are 1/3 of total income
  - Represents shift to new income path

\vfill
- **Size:** 4 million households treated
\vfill

- **Design:** Randomized rollout from 1998-2000
  - Afterwards, enrollment is non-random and based on **marginality index**
  - Two step: 1) village selected on index; 2) households selected in eligible villages

**Results Preview:** conditional cash transfer increases deforestation

## Data

- **Main variables:** village lat/lon, eligibility index, forest cover (satellite)
  - Enrollment by village until 2003
\vfill
- **Sample size:** 105,749 villages
\vfill
- **Unit of analysis:** only village centroids provided?
  - How do we estimate village-level impacts? Thiessen polygons
  - Area around a point where every location closer to the point than to all others
  - Problems?

## Thiessen/Voronoi Polygons
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ag2013_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Data Visualization
Eligibility cutoff: -1.2
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "49%", fig.align = "center"}

knitr::include_graphics(
                        path = c("assets/ag2013_fig2.png","assets/ag2013_fig3.png"),
                        auto_pdf = TRUE
                        )

```

## Empirical Strategy: OLS Equation

$$\Delta f_{i} = \alpha + \delta T_{i} + \beta'X_{i} + \epsilon_{i}$$

  - where $\Delta f_{i}$ is \% deforestation in polygon i over 2000-2003
  - $T_{i}=1$ if the locality enrolled by 2003
  - $X_{i}$ includes poverty

- But $T_{i}$ is not random!

\vfill

- If discontinuity is sharp, replace $T_{i}$ with $1[E_{i}>-1.2]$
- RDD: Sample window $-2 < I < -0.2$

## RDD Equation

- Discontinuity is fuzzy, not sharp
  - How do we know? Two reasons

- Solution: Use cutoff to instrument treatment probability

- Fuzzy RDD/IV:
$$\Delta f_{i} = \alpha + \delta T_{i} + \gamma I_{i} + \beta'X_{i} + \epsilon_{i}$$
$$\Delta T_{i} = \omega + \tau_{1}E_{i} + \tau_{2}E_{i}I_{i} + \tau_{3}M_{i} + \tau_{4}M_{i}I_{i} + \mu I_{i} + \Gamma' X_{i} + \epsilon_{i} $$ 
where $T_{i}$ is enrollment dummy, $E_{i}$ is eligibility cutoff dummy, $I_{i}$ is index, $M_{i}$ is dummy for region where enrollment increases rapidly

## Identification Assumptions

## Results: OLS and Sharp RDD
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ag2013_fig4.png",
                        auto_pdf = TRUE
                        )

```

## Results: Fuzzy RDD / Instrumental Variables
- Very strong first stage: eligiblity $\uparrow$ probability of enrollment ($\tau_{1}=0.8$)
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ag2013_fig5.png",
                        auto_pdf = TRUE
                        )

```

## Mechanisms

- RD results show that CCT increases deforestation
\vfill
- Why? What changes at the household level?
\vfill
- Approach 1: Use experimental sample (Progresa) with household survey data
  - Progresa had randomized rollout at village level for first 3 yrs

\vfill
- Approach 2: Heterogeneity by road density to study role of market access

## Mechanisms: Consumption channel
- Before (1997-1998) and after (2000) data on consumption
- 506 villages, 320 treated, 186 control (not-yet-treated)
- Treatment at the village level

Difference in differences:
$$ y_{it} = \gamma_{0} + \gamma_{1}Treat_{i} + \gamma_{2}Post_{t} + \gamma_{3} Treat_{i} \times Post_{t} + \epsilon_{it}$$

- $y_{it}$ = consumption
- $Treat_{i}=1$ if household $i$ in treated locality
- cluster at locality level bc randomization was at locality level
- How do we interpret $\gamma_{3}$?
\vfill
**Hypothesis: $\gamma_{3}>0$**

## Mechanisms: Market channel
- Demand shock must be met by supply, which drives land use change
  - e.g. $\uparrow$ consumption of milk met by $\uparrow$ in grazing land
\vfill

- But $\gamma_{3}$ captures partial equilibrium
  - Part of demand shock supplied locally. What about rest?
\vfill

- Hypothesis: If consumption is driving deforestation
  - Then effect should be larger when infrastructure quality is low
  - Low market access: demand met by local supply
  - High market access: demand shock propagates across markets

\vfill

## Triple Differences:

Heterogeneity by infrastructure quality:
\begin{align*} 
y_{it} & = \beta_{0} + \beta_{1}Treat_{i} + \beta_{2}Post_{t} + \beta_{3}(Treat_{i} \times Post_{t}) \\ & + \beta_{4}Road_{i} + \beta_{5} (Road_{i} \times Treat_{i}) + \beta_{6} (Road_{i} \times Post_{t}) \\ & + \beta_{7} (Road_{i} \times Treat_{i} \times Post_{t}) + \epsilon_{it}
\end{align*}

- $Road_{i}$ = inverse road density (km of road w/n 10km of village/polygon)

\vfill

**Hypothesis: $\beta_{7} > 0$**
\vfill

## Results: Increase in consumption of land intensive goods
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ag2013_fig6.png",
                        auto_pdf = TRUE
                        )

```
Higher beef and milk demand (land intensive products)

Demand-side impacts do not vary with market access (we did't expect it to)

## Results: No increase in local production
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ag2013_fig7.png",
                        auto_pdf = TRUE
                        )

```
Supply-side impacts do not vary with market access

## Results: Deforestation higher in places with poor market access
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ag2013_fig8.png",
                        auto_pdf = TRUE
                        )

```
\vfill
More deforestation in places with less market access (low road density)

Supports mechanism of increased demand for land-intensive goods

## Thoughts?

- Do we believe the story about increased demand for land-intensive goods?
  - Less market access means less access to inputs
  - Underinvest in technology, leading to higher deforestation
  - Maybe this is a story about technology, not consumption

- Can you think of an alternate explanation?

- Other concerns?

- This all points to shortcomings of RCTs in general
  - we cannot answer why

## Same results in Colombia! Malerba (2020)

- Question: What is the impacts of CCT on consumption, energy, and deforestation?

- Context: Familias en Accion CCT project (2001-2005)

- Design: Matched Difference in Differences

- Result: Increased beef and milk (land intensive) consumption
  - Mediated by markets
  - Negligible $\uparrow$ in deforestation (counterintuitive)
  - No impact on CO2

## Research Design
:::::::::::::: {.columns}
::: {.column width="45%"}
- CCT program launched in 2000
  - Non-random: Municipalities selected on amenities (banks, education, health)
  - 721 eligible: slow phase-in

\vspace{0.3cm}
- Study sample: 5,477 households
\vspace{0.3cm}
- Variation: compare outcome in treated hh before/after enrollment relative to control hh
\vspace{0.3cm}
- **Identification assumptions**

:::
::: {.column width="45%"}
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/malerba2020_fig2.png",
                        auto_pdf = TRUE
                        )

```
:::
::::::::::::::

## Results: Increased beed and milk consumption
Household DiD: $$ \Delta Y_{i} = \alpha + \delta T_{i} + X_{i} + \epsilon_{i} $$

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/malerba2020_fig1.png",
                        auto_pdf = TRUE
                        )

```
- Effect muted with low market access (very weak)
- Why is sample size declining from 5,477?

## Design: Environmental Impacts

- Unit of analysis: Municipality
\vfill
- Design: municipalities enrolled gradually over time
  - Many were not enrolled by 2004
\vfill
- Define $Treat_{i} = 1$ if enrolled by 2004
- $Post_{t} = 1$ after treatment, zero for 2000

$$y_{it} + \beta_{0} + \beta_{1} Treat_{i} + \beta_{2} Post_{t} + \beta_{3} (Treat_{i} \times Post_{t}) + \beta_{4} X_{it} + \epsilon_{it} $$
\vfill
- Variation: $\Delta$ forest b/w 2000/05 for municipalities enrolled by 2004 rel. to non-enrolled

## Results: No environmental impact
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/malerba2020_fig3.png",
                        auto_pdf = TRUE
                        )

```
- Weakly positive deforestation (counterintuitive)
- **Why?**

##

\centering
\huge Take a 5 minute break

## Back to the question
How does development affect the environment?
\vfill

- Income 
  - Land intensive consumption
  - **Energy**

\vfill
- Next time: Technology and Infrastructure

## Development, Energy, and the Environment

- Energy ladder: as income $\uparrow$, move from solid fuel $\rightarrow$ gas $\rightarrow$ electricity
  - Electricity may displace dirty energy (wood), which $\downarrow$ pollution

- But, electricity also increases total energy (fridge, AC)
  - ...which $\uparrow$ pollution
  - energy footprint increases with income

- Unless generated from solar, wind, etc

- People also buy cars as they get richer...

## GDP and Car Ownership
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "40%", fig.align = "center"}

knitr::include_graphics(
                        path = c("assets/ekc_cars.png","assets/ekc_legend.png"),
                        auto_pdf = TRUE
                        )

```
Of course, these are correlations

## But first, some theory

- Imagine we can exogenously vary income to poor households
\vfill
- Will transfers to initially poor households yield **lumpy** investments?
  - Movements up the ladder involve big purchases
  - May explain why 1.3 billion people lack electricity (Gertler et. al, 2016) 
\vfill
- Credit constraints play an important role
  - Without credit markets, buy now but have nothing in next period
  - With savings, can save now, buy later (smooth consumption)
  - With borrowing, buy now, smooth consumption over time
\vfill
- Lets formalize this

## Model Set up

- Two periods with no discounting
- Agent $i$ can consume two goods
  - non-durable gives per period utility: $u'(\cdot)>0$, $u''(\cdot)<0$
  - lumpy durable that gives static per period utility R, if owned

- Durable price = P; non-durable is numeraire

- Let $Y_{1}$, $Y_{2}$ be per period 1, 2 income

- Define $\bar{Y} = 1/2(Y_{1}+Y_{2})$ = average per period income

## No credit constraints

Without credit constraints, if don't buy, total utility is:
$$u(\bar{Y}) + u(\bar{Y}) = 2u(\bar{Y})$$
If buy durable, spend $P$ and spread cost equally across periods:
$$2u(\bar{Y}-\frac{P}{2}) + 2R$$
Purchase durable iff:
$$ u(\bar{Y}) - u(\bar{Y}-\frac{P}{2}) \leq R$$
**$u''(\cdot)<0$ implies acquisition increases in income**

## 1) With credit constraints: Buy in period 1

If don't purchase, total utility is:
$$ u(Y_{1}) + u(Y_{2}) $$

If purchase in period 1: 
$$ u(Y_{1} - P) + 2R + u(Y_{2}) $$

Purchase in period 1 iff:

$$ \frac{u(Y_{1}) - u(Y_{1}-P)}{2} \leq R $$

## 2) With credit constraints: Wait to buy in period 2

If don't purchase, total utility is: 
$$ u(Y_{1}) + u(Y_{2}) $$
If save in period 1, and buy in period 2:
$$ 2u(\bar{Y} - \frac{P}{2}) + R$$
Wait to buy in period 2 if:
$$u(Y_{1})+u(Y_{2}) - 2u(\bar{Y} - \frac{P}{2}) \leq R $$

## Intuition of Model

- $u''(\cdot)<0$ means consumers gain from smoothing consumption

- Unconstrained household will buy in period 1
  - Use period 2 income (loan) to smooth consumption

- Credit constrained households cannot do this
  - If buy now, magnify consumption inequality across periods
  - Or, wait and buy in period 2, but delay utility gain

- Higher income in any period leads savers to buy more

- But, first period buyers respond only to first period income (lemma 1)

- But increase in *cumulative* income increases buying through delay/saving (lemma 2)

## Testable Predictions

- Prediction 0: Acquisition increasing in income
- Prediction 1: S-shaped curve in acquisition
<!-- - Prediction 2: Faster income growth leads to more period 2 adoption -->

## Gertler et al. (2016): Demand for energy-using assets among middle class
:::::::::::::: {.columns}
::: {.column width="45%"}
- **Question:** How does income affect energy demand?
\vspace{0.3cm}
- **Motivation:** 1.3 billion people live without electricity
\vspace{0.3cm}
- Should we expect linear climb up the energy ladder as incomes rise?
  - what about credit markets?

:::
::: {.column width="45%"}
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gert2016_fig1.png",
                        auto_pdf = TRUE
                        )

```
:::
::::::::::::::

## Non-linear relationship between income and fridges
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "65%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gert2016_fig2.png",
                        auto_pdf = TRUE
                        )

```
- Are fridges representative of energy-intensive assets?
- Can aggregating non-linearities $\rightarrow$ linearity?

## S-shape appears robust
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "50%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gert2016_fig3.png",
                        auto_pdf = TRUE
                        )

```

## Empirical Setting: Mexico Oportunidades (again)

- 320 randomly selected communities given early treatment (April 1998)
  - 52\% were ``eligible'' for cash transfer
\vfill
- 186 randomly selected communities given late treatment (October 1999)
\vfill
- This paper: N=506 communities, 10,000 households surveyed (1997-2007)
  - Surveys done in  1998, 1999, 2000, 2003, and 2007
\vfill
- Compare early and late treatment households (**not treatment vs. control**)
  - Similar b/c both groups selected on ``vulnerability'' characteristics
  - Strongly balanced on covariates
  - Assumes timing is random

## Empirical Design

- **Prediction 0:** Probability of asset purchase $\uparrow$ in income

$$h(a_{it}) = Pr(a_{it}=1 | a_{it-1}=0) = \gamma_{0} + \alpha_{1} CI_{it} + \beta X_{i} + \beta_{t}F_{i} + R_{rt} + \epsilon_{it}$$

  - where h(a) is the prob. that $i$ buys $a$ in time $t$ conditional on not having it in $t-1$. 
  - $CI_{it}$ is cumulative income
  - $X_{i}$ includes baseline household controls
  - $F_{i}$ are family structure variables interacted w/ survey round
  - $R_{rt}$ are state-year fixed effects

Hypothesis: $\alpha_{1} >0$

## Empirical Design

- **Prediction 1:** S-shaped curve in acquisition
$$h(a_{it}) = Pr(a_{it}=1 | a_{it-1}=0) = \gamma_{0} + \alpha_{1} CI_{it} + \alpha_{2} CI^{2}_{it} + \beta X_{i} + R_{rt} + \epsilon_{it}$$

Hypothesis: $\alpha_{2} \lessgtr 0?$

<!-- ## Empirical Design -->

<!-- - **Prediction 2:** conditional on having same level of cumulative income, households which accumulated income slower are less likely to acquire asset -->

<!-- $$h(a_{it}) = Pr(a_{it}=1 | a_{it-1}=0) = \gamma_{0} + \alpha_{1} CI_{it} + \alpha_{3} Early_{i} + \alpha_{4} (Early_{i} \times CI_{it}) + \beta X_{i} + R_{rt} + \epsilon_{it}$$ -->

<!-- - where $Early_{i}=1$ if $i$ began receiving transfers 18 months before control households -->
<!-- - $Early_{i}$ is synonymous with treated, and $Late_{i}$ is control (not-yet-treated) -->

## Measurement issues

- Cumulative Income is sum of wage, farm, business income, CCT transfers
  - Reporting biases in wage data
  - Income reported in survey year, not cumulative
  - Authors interpolate across years, adding noise
\vfill
- Double counting problem if households invest CCT money into business
\vfill
- **Solution:** use potential CCT **transfers** to instrument cumulative **income**

## Identifying Variation

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gert2016_fig4.png",
                        auto_pdf = TRUE
                        )

```

Use maximum cumulative transfer as **instrument** for actual (cumulative) transfer

Plus variation from early/late enrollment

## Identification Assumptions

- What is the exclusion restriction?
- Is it reasonable?

## Results: Prediction 0
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gert2016_fig5.png",
                        auto_pdf = TRUE
                        )

```
- For every 10,000 pesos, probability of buy fridge $\uparrow$ by 4.7 pp.

## Results: Prediction 1
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gert2016_fig6.png",
                        auto_pdf = TRUE
                        )

```
- $\alpha_{2}>0$ implies convex relationship b/w cumulative income and asset purchase
- Since all households are poor, don't expect any to be past second inflexion point

## Thoughts on paper?

## Research gaps

- Does earned income have different effect than transfers?
\vfill
- Can we trace out the full S-curve?
\vfill
- Is there an S-curve for other technologies?
\vfill
- Which externalities have highest income elasticities?
\vfill

## Next week

- How does development affect the environment?
  - **Technology and Infrastructure**

##

\centering
\Huge Lecture 3

\large How does development affect the environment?

## Housekeeping

- Research proposal first draft: Oct 3rd
  - If you find idea w/n 10 days, gives 2 weeks to write
\vfill
- Class presentations: separate Canvas upload each week
\vfill
- Replication project?

## Recap: Alix-Garcia et al. (2013)

- Do we believe the story about increased demand for land-intensive goods?
  - Less market access means less access to inputs
  - Underinvest in technology, leading to higher deforestation
  - Maybe this is a story about technology, not consumption

- Can you think of an alternate explanation?
  - Talk to your neighbor for five minutes

- Other concerns?

## Recap

\vfill
- Higher income leads to land-intensive **consumption**
  - Beef, milk, etc.
  - This requires clearing land for agriculture
  - Effect depends on market access
\vfill
- Higher income leads to energy-intensive **consumption**
  - But the relationship is non-linear
  - Role of credit constraints

## Today

- Development increases access to **productive** capital
  - How do we think about environmental implications?
\vfill
- Case I: Development increases agricultural productivity
  - Abman et al. (2023)
\vfill
- Case II: Development spurs infrastructure development
  - Garg et al. (2023)

## Technology and infrastructure

- As economies develop, technology and infrastructure improve
\vfill
- Productivity boost means firms produce at lower cost
  - Outward shift in supply curve $\rightarrow$ \textcolor{brown}{bad for environment}
  - But, technological progress reduces resource intensity $\rightarrow$ \textcolor{green}{good for environment}
\vfill
- Most active areas of research:
  - agricultural productivity
  - infrastructure

## Agricultural productivity and the environment

- **Boserup's hypothesis **: extensive margin
  - improvements in land productivity $\uparrow$ ag land value
  - leads to pressure on forests to clear land for ag
\vfill
- **Borluag's hypothesis**: intensive margin
  - *Under factor market constraints*, $\rightarrow$ productivity spurs intensification
  - ``spares'' land for nature and puts less pressure on forests
\vfill
- **Jevon's paradox**: general equilibrium
  - technology for increasing yield also lowers MC of producing given amount
  - Ag supply curve shifts out and may dominate gains in yield
  - Leads to net increase in land demand

**Takeaway: this is an empirical question!**

## Existing Work

- Theory: (Angelsen and Kaimowitz,2001; Balsdon, 2007; Goldstein et al., 2012; Green et al., 2005; Phalan et al.,2016; Takasaki,2006). 
  - Most models yield ambiguity
\vfill
- Most work on agricultural productivity and deforestation
  - Intensive margin dominates: Abman et al. (2024), Abman and Carney (2020a), Abman and Carney (2020b), Assuncao et al. (2023)
\vfill
- New work: market access and deforestation 
  - Abman and Lundberg (2024)
\vfill
- Open Topics:
  - Effect of agriculture on deforestation in the first place? (Madhok et al. 2024)
  - Role of factor markets? Labor vs. input markets
  - Mechanisms behind market access effects
  - Agriculture, migration and deforestation

## Conceptual framework

- Continuum of agents with heterogeneous outside options
  <!-- - Changes in returns to ag affect number of agents who select into farming -->

- Farmers allocate labor b/w farming on existing land, or cultivate new land

- Let $\gamma$ denote household labor allocated to new land

- Production function for new and existing land are $f(\cdot)$ and $g(\cdot)$

- Exogenous productivity shock, $\Omega$, improves productivity of existing land more

$$ \frac{\partial g(l)}{\partial \Omega} \geq \frac{\partial f(l)}{\partial \Omega} \geq 0 \quad \quad \forall l$$

## Conceptual framework

- Farmers are factor market constrained (Conning and Udry, 2007)
  - Cannot hire labor from market

- Total deforestation given by new land cleared for agriculture:
$$ D^{*} = n^{*} \cdot \gamma^{*} $$

- Differentiating with respect to $\Omega$ gives:
$$\frac{\partial D^{*}}{\partial \Omega} = \underbrace{n^{*}\frac{\partial \gamma^{*}}{\partial \Omega}}_\text{$\lessgtr 0$} + \underbrace{\gamma^{*}\frac{\partial n^{*}}{\partial \Omega}}_\text{$\lessgtr 0$}$$

- What does each term represent and what is the sign?

## Abman et al. (2023): Agricultural Productivity and Deforestation

- **Goal:** To test $\frac{\partial D^{*}}{\partial \Omega}$, we need random variation in $\Omega$

- **Question:** what is the impact of ag productivity on deforestation?

- **Context:** Large scale ag extension program in Uganda (2008-2013)

- **Design:** Leverage Spatial discontinuity in village eligibility

- **Results:** improvement in agricultural productivity; reduction in deforestation

## Background: Why Uganda?

- Forest cover shrunk from 24\% to 9\% from 1990-2015

- Most forest land is **privately owned**
  - Land owners allowed to covert forest (Land Act, 2010)
  - Cannot use state powers to limit deforestation

- Need an indirect mechanism to incentivize conservation

## Background: BRAC Extension Program

- Launched in 2008

- Aims to extend modern cultivation techniques to smallholders

- Two step treatment:
  - Training: ``model farmers'' trained in modern cropping, irrigation, weeding, pest control. Then, set up demo plot and train others
  - Extension: Promoters selected to sell subsidized HYV seeds in their villages

- **Key feature: Program limited to villages within 6km of each BRAC branch**

## Data

- Deforestation from GFC (Hansen et al., 2013)
  - Pro: high resolution (30m)
  - Con: 1) only captures loss not gain, 2) everything relative to 2000

- Lat/lon of villages within 12km of BRAC office (N=807 villages)

- Attribute forest pixel to village if w/n 400m of village centroid
  - Rationale: median household distance to village center = 400m
  - Problems?

- BRAC agricultural survey on 7,781 households (451 villages) in 2011
  - Detailed cropping practices from last two seasons

## Empirical Design: Spatial Regression Discontinuity

- Estimate IIT since we do not know true treatment designation
\vfill
- Local linear regression to estimate left and right side of discontinuity

$$ \beta = lim_{z \uparrow 0} E[Y|z_{i} = z] - lim_{z \downarrow 0} E[Y|z_{i} = z]$$

- where running variable, $z_{i}$, is distance of village from 6km cutoff
  - $z \leq 0$ means village had access to extension
\vfill
- For estimation, use 2km bandwidth
  - Higher regression weights on points nearer to cutoff (triangle kernel)

## Aside: RD with local linear regressions

- Left side of cutoff
$$ Y_{i} = \alpha_{L} + \beta_{L}(z_{i}-c), \quad \quad \text{for $z_{i} \leq$ c} $$
- Right side of cutoff
$$ Y_{i} = \alpha_{R} + \beta_{R}(z_{i}-c), \quad \quad \text{for $z_{i} \geq$ c} $$
- Treatment effect at threshold:
$$ \beta = \alpha_{R} - \alpha_{L}$$
since $\alpha$ are the intercepts on each side of cutoff

## Identification Assumptions?

<!-- eligibility uncorrelated with unobserved factors that drive deforestation and vary at the cutoff -->

<!-- what if baseline forest cover increases around 6km, and more degraded areas selected for program? -->

## Identification Assumptions
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/abman2020_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Identification Assumptions
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/abman2020_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Results: 14\% less deforestation in treated villages
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/abman2020_fig3.png",
                        auto_pdf = TRUE
                        )

```

## Mechanisms: Intensification
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "97%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/abman2020_fig4.png",
                        auto_pdf = TRUE
                        )

```
- Techniques to address nutrient depletion **on existing land**
- No evidence of increased fertilizer, pesticides, or HYV seed adoption

## Intensification vs. Extensification
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/abman2020_fig5.png",
                        auto_pdf = TRUE
                        )

```
- Revenue/acre is a proxy for yields
- No extensification $\rightarrow$ binding constraint on land clearing

## Discussion

- Empirical issues
- Alternative explanations
- External validity

## Back to the Question
How does development affect the environment?
\vfill

- Income
  - Land intensive consumption
  - Energy

\vfill
- Access to capital
  - Technology
  - **Infrastructure**

## Infrastructure and the Environment
- Roads can increase deforestation by:
  - $\downarrow$ MC (transport costs) of forest products, increasing supply
  - $\uparrow$ market access for agricultural products, causing extensification
  - $\uparrow$ land value, leading to agglomeration effects
\vfill
- Roads can increase forest cover by:
  - $\uparrow$ access to forest product substitutes (i.e. clean fuel vs. firewood)
  - $\uparrow$ access to external labor/capital markets, lowering need for land-clearing
\vfill
- Net effect is ambiguous

## What about Infrastructure and Air Pollution?

- Lets build a conceptual framework
- Talk to your neighbor for 10 mins

## Garg et al. (2023): Rural Roads, Labor Exits, and Crop Fires

- **Question:** what is the impact of roads on air quality?
  - what is the role of labor reallocation as a mechanism?

- **Context:** India builds last-mile roads using population cutoff

- **Design:** Regression discontinuity with crop fires as outcome

- **Results:** Roads $\uparrow$ crop fires and PM2.5 levels in a village

- **Mechanism:** Roads $\rightarrow$ labor exit $\rightarrow$ higher wages $\rightarrow$ labor-saving technology

## Background

- Labor is misallocated in agriculture in developing countries
  - $MP_{labor}$ lower in agriculture than other sectors

- Sources of misallocation is a big question in dev econ

- Many policies aimed at reducing misallocation
  - e.g. road-building (improve market access)

- Do these have externalities?
  - **innovation**: where do the externalities come from?

## Context: India

- 14 of the 20 most polluted cities in the world are in India

- Important pollution source: crop burning

- Pros:
  - Clear rice straw residue \& undergrowth b/w cropping seasons
  - Cheaper than labor/capital alternative

- Cons:
  - Carries civil/criminal penalty (weakly enforced)
  - Reduces soil nutrients and decreases ag productivity
  - Damages human health

- Roads may facilitate agricultural labor loss
  - Will farmers turn to crop burning, despite the costs?

## Research Design

- PMGSY aims to provide all-weather roads to all unconnected villages in India
\vfill
- Prioritize larger villages first
  - Connect villages with population > 1000 by 2003
  - Connect villages with population > 500 by 2007
  - Connect villages with population > 250 afterwards
\vfill
- State-by-state basis
  - If connected all large villages, can proceed to smaller ones immediately

## Data
- Roads: SHRUG database 
  - N=11,151 villages w/o paved road at baseline (2001)
  - Main variable: road construction date in village
\vfill
- Fires: NASA EODIS
  - \# fires w/n 10km of village polygon
\vfill
- Pollution: Van donkelaar et al. (2016) measure PM2.5 at 10km resolution
  - Black Carbon and Organic Carbon from biomass/non-biomass sources
\vfill
- Infant mortality: NFHS-IV (2015-16)
  - child-birthyear panel from 2001-2013
  - Identify survey clusters (villages) within 50km of each PMGSY village

## Empirical Strategy: Fuzzy RDD
First Stage:
\begin{align*}
  Road_{vdst} & = \gamma_{0} + \gamma_{1} 1(pop_{vds} \geq T) + \gamma_{2}(pop_{vds} - T) \\ & + \gamma_{3}(pop_{vds}-T) \times 1(pop_{vds} \geq T) + \theta X_{vds} + \mu_{d,h} + \rho_{t} + \epsilon_{vdst}
\end{align*}

Second Stage:
\begin{align*}
  Y_{vdst} & = \beta_{0} + \beta_{1} Roads_{vdst} + \beta_{2}(pop_{vds} - T) \\ & + \beta_{3}(pop_{vds}-T) \times 1(pop_{vds} \geq T) + \delta X_{vds} + \eta_{d,h} + \omega_{t} + \epsilon_{vdst}
\end{align*}

- $Road_{vdst}=1$ if village $v$ received road by 2013
- $Y_{vdst}$ is outcome (fires, pollution, mortality)
- $pop_{vds}$ is village population in 2001
- $T$ is treatment threshold (500 or 1000, depending on state)

## Covariate Balance
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/garg2023_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Results
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/garg2023_fig2.png",
                        auto_pdf = TRUE
                        )

```
- Note: optimal bandwidth = +/- 84 around population threshold
- Road access increases annual crop fires by 60\%
- Road access increase PM2.5 level by 0.5 $\mu g/m^{3}$ (1.1\%)

## Does pollution result also capture increased driving on new roads?
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/garg2023_fig3.png",
                        auto_pdf = TRUE
                        )

```

## Mechanism: Roads increase ag wage rate
- Data: REDS Survey (1999 and 2006)
- Design: Diff-in-Diff (221 villages, 100 districts, 17 states)
  - Treat $=1$ if $v$ received road b/w 1999-2006; Post=1 in 2006
$$ Y_{vt} = \alpha + \beta_{1}(Treat_{v} \times Post_{t}) + \gamma_{v} + \theta_{t} + \epsilon_{vt}$$
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/garg2023_fig4.png",
                        auto_pdf = TRUE
                        )

```

## Mechanism: Roads induce labor exit
- Back to RDD design
- Data from Socioeconomic and Caste Census (2013), provided in SHRUG
- Outcome: ag. labor share in districts with high/low ag wage
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/garg2023_fig6.png",
                        auto_pdf = TRUE
                        )

```
- More labor exit in districts with lower baseline ag wages

## Mechanism: places with more labor exit burn more biomass
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/garg2023_fig5.png",
                        auto_pdf = TRUE
                        )

```
- Fire/pollution impact greater in places with lower baseline ag wage
- Road generates more labor exit in these places

## Discussion

- Do you believe the story?
- Alternative explanations

## Next Time

- Presentations
- How does the environment affect development?
  - **Health**
  - Productivity