### 2024-11-11 lectures/module4/module4.Rmd

---
title: "| Environmental and Development Economics\n| Module 4 - WTP for Environmental Quality
  \n"
author: |
  | Raahil Madhok
  | UMN Applied Economics
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Montpellier
  slidy_presentation: default
  ioslides_presentation: default
classoption: aspectratio=169
header-includes:
- "\\AtBeginDocument{\\title[Module 4 - WTP for environment]{Environmental and Development
  Economics \\\\ Module 4 - WTP for Environmental Quality in LMICs}}"
- \input{header.tex}
- \usepackage{gensymb}
- \usepackage{cancel}
---

## 
\centering
\Huge Lecture 6

\large WTP Estimation: Revealed Preference Models

## Housekeeping

- Research Proposal: you do NOT have to carry out analysis
\vfill
- First draft: Oct 3rd
\vfill
- Replication: Oct 21st
  - Start soon

## Today

- **Guiding question:** what is the WTP for environmental quality in LMICs?
\vfill
- Today's focus: How do we even measure this?
  - Behavioral models
  - Choice probabilities
  - Estimation
  - Application to WTP

\vfill
- Next time: why is WTP so low in developing countries?

## Estimating Non-market Value of Goods and Services

- **Stated Preference:**
  - Hypothetical data to estimate **ex-ante** WTP
  - Contingent valuation, contingent behavior
  - Respondents directly asked about WTP (phone, mail, etc.)
  - Issue: hypothetical bias and strategic bias
\vfill

- **Revealed preferences**
  - Behavioral data to estimate **ex-post** WTP
  - Travel cost, averting behavior, hedonic price
  - Pro: based on **actual** choices
  - Con (?): need a **behavioral model** in which to analyze choices

## Behavioral Models

- Our goal is to define a behavioral process generating agent's choice

- Let agent face choice among set of options (products, actions, etc.)
  - What leads agent to choose one?

- Let some factors ($x$) be observed by researcher, and some not ($\epsilon$)

- Factors relate to agent choice through a **behavioral process**:

$$y = h(x, \epsilon)$$

- \(h(x, \epsilon)\) determines the selected choice based on these factors

## Probabilistic Choice
- Since \(\epsilon\) unobserved, the outcome is probabilistic

- **Choice probability** of choosing $y$ derived by assuming $\epsilon \sim f(\epsilon)$
$$P(y | x) = \text{Prob}(\epsilon \quad \text{s.t.} \quad h(x, \epsilon) = y)$$

- This probability is much more tractable i.e., fully characterized by $f(\epsilon)$

- We can rewrite this as the expected value of an indicator function:
$$P(y | x) = \int I[h(x, \epsilon) = y] f(\epsilon) d\epsilon$$

- where $I[h(x, \epsilon) = y] =1$ when statement in brackets is true
- Integral of behavioral process indicator over all possible values of unobservables

## Evaluating the Choice Probability

- Method 1: Closed-form solution
  - For certain specifications of $h$ and $f$, $P(y|x)$ calculated via a formula
  - Main example is Logit

\vfill
- Method 2: Simulation
  - Some specifications of $h$ and $f$ have no closed form solution
  - We simulate $P(y|x)$ by taking random draws from $f$ and taking the average

## Evaluating the Choice Probability: Closed Form

**First, define the behavioral model**

- Consider a binary model where agent considers whether to take an action
- Agent gets utility (+/-) from the action:

$$U= \beta'x + \epsilon$$

  - $U$: Utility of taking an action; $\beta' x$: Observed component of utility.
  - $\epsilon$: Unobserved component; lets assume **logistic distribution**.

- Take action if utility is positive:
$$P = \int I[\beta' x + \epsilon > 0] f(\epsilon) d\epsilon$$

## Closed Form Solution
\begin{align*}
P & = \int I[\beta' x + \epsilon > 0] f(\epsilon) d\epsilon \\
 & = \int_{\epsilon > -\beta' x} f(\epsilon) d\epsilon \\
 & = 1 - F(-\beta' x) \\
 & = 1 - \frac{1}{1 + e^{\beta' x}} \\
 & = \frac{e^{\beta' x}}{1 + e^{\beta' x}} \\
\end{align*}

## Evaluating the Choice Probability: Simulation

Recall: $P(y | x) = \int I[h(x, \epsilon) = y] f(\epsilon) d\epsilon$

\vfill

- **Step 1:** Take draw of $\epsilon$ from $f(\epsilon)$
  - Label it $\epsilon^{1}$, denoting first draw
\vfill
- **Step 2:** Check whether $h(x, \epsilon^1)=y$
  - If so, create $I^{1}=1$, otherwise $I^{1}=0$
\vfill
- **Step 3:** Repeat this R times and collect $I^{r}$ for $r=1,...,R$
\vfill
- **Step 4:** Average **simulated** probability: $P(i | x) = \frac{1}{R}\sum_{r=1}^{R} I^{r}$
  - Proportion of times that draws of $\epsilon$, when combined with $x$, yield $y$

## 
\huge Lets define a realistic behavioural model

\large Grounded in economic theory

## Random Utility Model

**Behavioural model:**

- Agent, $n$ faces choice among $J$ alternatives
  - Choice set must be 1) mutually exclusive, 2) exhaustive, 3) finite

- Utility from $j$ is: $U_{nj}$ for $j=1,...,J$

- Utility known to agent but not the researcher

- Agent chooses $j$ that maximizes utility:

$$ \text{Choose i iff} \quad U_{ni} > U_{nj} \quad \forall j\neq i$$

## Random Utility Model

**The Researcher**

- Observes attributes of alternatives, $x_{nj} \quad \forall j$

- Observed attributes of agent, $s_{n}$

- Let $V_{nj} = V(x_{nj}, s_{n}) \quad \forall j$ (representative utility)

- **V can depend on parameters unknown to researcher**
  - But can be structurally estimated...

## Random Utility Model

- Since $V_{nj} \neq U_{nj}$, we can write:
$$U_{nj} = V_{nj} + \epsilon_{nj}$$

  - where joint density of random vector $\epsilon_{n} = (\epsilon_{n1},...,\epsilon_{nJ}) = f(\epsilon_{n})$

- Then the probability than agent $n$ chooses alternative $i$ is:
\begin{align*}
P_{ni} & = P(U_{ni} > U_{nj}, \forall j \neq i) \\
 & = P(V_{ni} + \epsilon_{ni} > V_{nj} + \epsilon_{nj}, \forall j \neq i) \\
& = P(\epsilon_{nj} - \epsilon_{ni} < V_{ni} - V_{nj}, \forall j \neq i) \leftarrow \text{{\color{blue} CDF!}} \\
 & = \int I(\epsilon_{nj} - \epsilon_{ni} < V_{ni} - V_{nj}) f(\epsilon) d\epsilon
\end{align*}
- Different DCMs obtained from different choices of $f(\epsilon)$

## Logit

- Assumes $\epsilon_{nj}$ distributed iid extreme value
$$ \text{PDF:} \quad f(\epsilon_{nj}) = e^{-\epsilon_{nj}} e^{-e^{-\epsilon_{nj}}}, \quad \text{CDF:} \quad F(\epsilon_{nj}) = 1 - e^{-e^{\epsilon_{nj}}}$$
- Note: if $\epsilon$'s are iid EV, then $\epsilon_{nji}^{*} = \epsilon_{nj} - \epsilon_{ni}$ is distributed logistic:
$$F(\epsilon_{nji}^{*}) = \frac{e^{\epsilon_{nji}^{*}}}{1+{\epsilon_{nji}^{*}}}$$

## Aside: Extreme Value Distribution

:::::::::::::: {.columns}
::: {.column width="45%"}

**Useful Properties**

\vspace{0.3cm}

1. Mathematical tractability
    - Closed form choice probability

\vspace{0.3cm}
2. Long right tail
    - Unobservables can drive unexpected choices

\vspace{0.3cm}

3. IIA property

\vspace{0.3cm}

4. Connection to logistic regression

\vspace{0.3cm}

5. Easy to (structurally) estimate

:::
::: {.column width="45%"}
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ev_pdf.png",
                        auto_pdf = TRUE
                        )

```
:::
::::::::::::::

## Logit

- The probability that agent $n$ chooses alternative $i$:
\begin{align*}
P_{ni} & = P(V_{ni} + \epsilon_{ni} > V_{nj} + \epsilon_{nj} \quad \forall j\neq i) \\
&= P(\epsilon_{nj} < \epsilon_{ni} + V_{ni} - V_{nj} \quad \forall j\neq i) \leftarrow \text{{\color{blue} invoke iid assumption!}} \\
 &= \int \left( \prod_{j \neq i} e^{-e^{-(\epsilon_{ni} + V_{ni} - V_{nj})}} \right) e^{-\epsilon_{ni}}e^{-e^{-\epsilon_{ni}}}d\epsilon_{ni} \\
&= \frac{e^{V_{ni}}}{\sum_{j}e^{V_{nj}}}
\end{align*}
\vfill
- **Preview:** If V is linear in parameters, $V_{nj} = \beta'x_{nj}$, then: $P_{ni} = \frac{e^{\beta'x_{ni}}}{\sum_{j}e^{\beta'x_{nj}}}$

## Independence of Irrelevant Alternatives

- Crucial assumption in logit model

- Relative odds of choosing b/w 2 $i$ and $k$ does not change when $j$ introduced

- For any two alternatives $i$ and $k$, the ratio of logit probabilities is:
\begin{align*}
\frac{P_{ni}}{P_{nk}} &= \frac{e^{V_{ni}}/ \sum_{j}e^{V_{nj}}}{e^{V_{nk}}/ \sum_{j}e^{V_{nj}}} \\
&= \frac{e^{V_{ni}}}{e^{V_{nk}}} = e^{V_{ni}-V_{nk}}
\end{align*}

- Ratio does not depend on any alternatives other than $i$ and $k$!
  - Denominators cancel

## Discussion of IIA Assumption

- Advantages
\vfill

- Disadvantages

##
\centering
\huge 5 minute break

## Estimation: Maximum Likelihood Structural Estimation

- Suppose we have have data on $N$ agent's choices
- The likelihood of observing the choices made by each individual is:

$$L(\beta) = \prod_{n=1}^{N} \prod_{i} (P_{ni})^{y_{ni}}$$

- where \(y_{ni} = 1\) if alternative \(i\) is chosen, and 0 otherwise
- Log-likelihood is easier to maximize:
$$LL(\beta) = \sum_{n=1}^{N} \sum_{i} y_{ni} \log(P_{ni})$$

- ML estimator is the $\beta$ that maximizes this function

## Method of Moments

- FOC of $LL(\beta)$ $\leftrightarrow$ $\frac{\partial LL(\beta)}{\partial \beta}=0$
- This can be rewritten as:
$$\frac{1}{N} \sum_{n=1}^{N} \sum_{i} y_{ni} x_{ni} = \frac{1}{N} \sum_{n=1}^{N} \sum_{i} P_{ni} x_{ni}$$

- LHS: mean $x$ over choices by sampled agents
- RHS: mean $x$ over **predicted** choices by sampled agents

- $\beta$ found s.t. predicted mean of each explanatory variable equals observed mean
  - Model reproduces observed averages in the data
  - Also known as **matching on the first moment**

## Limitations

- Cannot represent random taste variation

- Implies proportional substitution across alternatives (IIA)

##
\Large What if we want to estimate more moments of $\beta$?

## Specifying Taste Variation

- Supposed agent now has the utility: $U_{nj} = {\color{red} \beta_{n}}'x_{nj} + \epsilon_{nj}$
  - where $\beta_{n}\sim f_{n}(\beta|\Theta)$ and  $\epsilon_{nj}$ still iid EV

- If we observed $\beta_{n}$, then we are back to logit: $P_{ni}(\beta_{n})=\frac{e^{\beta_{n}'x_{ni}}}{\sum_{j}e^{\beta_{n}'x_{nj}}}$

- Instead, we integrate logit over all possible parameters of $\beta_{n}$:

$$P_{ni} = \int \underbrace{\left(\frac{e^{\beta_{n}'x_{ni}}}{\sum_{j}e^{\beta_{n}'x_{nj}}}\right)}_\text{logit probability} \underbrace{f(\beta | \Theta)d(\beta)}_\text{mixing distribution} \quad \leftarrow \text{{\color{blue} no closed form solution!}}$$

- Mixing distribution usually normal: $f(\beta|\Theta) = f(\beta|\mu,\sigma)$
- Note: $P_{ni}$ is a function of $\Theta$, not $\beta$! ($\beta$ is integrated out)

## Aside: Properties of Mixed Logit

- Relaxes IIA assumption!
\vfill
- Ratio of mixed logit probabilities, $P_{ni}/P_{nj}$ depends on all data
  - Denominators are inside the integral and do not cancel
\vfill
- Enables correlation in unobserved factors across alternatives
  - Flexible substitution patterns

## Estimation: Simulation

**Steps for Simulating Mixed Logit Probability**

1. Draw a value $\beta^r$ from $f(\beta | \Theta)$, where $r$ is which draw
2. Calculate the logit formula $P_{ni}(\beta^r)$ using the drawn value
3. Repeat steps 1 and 2 for *R* draws.
4. Average the results to obtain the simulated probability:

$$ \hat{P}_{ni}= \frac{1}{R} \sum_{r=1}^{R} P_{ni}(\beta^r)$$

## Simulated Log-Likelihood Estimation

- As before, insert simulated $\hat{P}_{ni}$ into log-likelihood:

$$SLL = \sum_{n=1}^{N} \sum_{j=1}^{J} y_{nj} \ln \hat{P}_{ni}$$

- where $y_{nj} = 1$ if $n$ chose $j$ and zero otherwise

- Maximum simulated likelihood estimator (MSLE) is the $\Theta$ that maximizes SLL
\vfill
**This recovers taste distribution $(\mu, \sigma)$ that matches observed choices to predicted choice probabilities as close as possible**

##
\centering
\huge Lets do a simple application

\large What is the WTP for biodiversity?

## WTP for Species Diversity in India

- **Context:** people travel to ``hotspots'' to see pretty birds
  - Birds are a good proxy for overall biodiversity
\vfill
- **Data:** individual level data from citizen science app (eBird)
  - location, alternative sites, income, **travel cost**, **species richness**
\vfill
- **Behavioral model:** choose site that maximizes utility
\vfill
- **Question:** what cost will you incur to $\uparrow$ utility from biodiversity?
  - How does your ``price'' change for an additional unit of biodiversity? (MWTP)

## Deriving WTP in Practice

- Agent $i$'s utility from site $j$ at time $t$ is:
\begin{align*}
U_{ijt} & = V_{ijt} + \epsilon_{ijt} \\
& = \underbrace{\beta_{1}}_{\text{MU Income}}(y_{i} - c_{ijt}) + \underbrace{(\beta_{2} + \eta_{i})}_{\text{MU EQ}} e_{jt} + ... + \epsilon_{ijt}
\end{align*}

- If stay home: $U_{i0t}=\beta_{1}y_{i} + \epsilon_{i0t}$
- Cost s.t. $i$ indifferent b/w travelling and avoiding cost:
$$c_{ijt}* = \frac{1}{\beta_{1}} [(\beta_{2} + \eta_{i})e_{jt} + \beta_{3} x_{ijt} + ... + (\epsilon_{ijt}-\epsilon_{i0t})]$$
- MWTP for one more species, assuming $\eta_{i}\sim N(0,\sigma)$: 
$$MWTP=\boxed{\frac{\partial c_{ijt}*}{\partial d_{jt}} = \frac{\hat{\beta_{2}}}{\hat{\beta_{1}}} + \xcancel{\frac{\hat{\eta_{i}}}{\hat{\beta_{1}}}}}$$

## Implementation

- Data
  - attributes of choice and alternatives for each **choice occasion**
  - Variable *choice=1* if chosen, otherwise 0
  - Data can get huge, and simulation takes forever
  - solution: restrict choice set
\vfill
- Stata
  - clogit: conditional logit (fast)
  - mixlogit: mixed logit estimation (slow)
\vfill
- R
  - *mlogit* library
  - Exercises and applications: https://cran.r-project.org/web/packages/mlogit/vignettes/c5.mxl.html

## Aside

- **If you want to get involved in this project, let me know**

- Goal: estimate WTP for species diversity in India

##
\centering
\huge Lecture 7

\large Why is WTP for Environmental Quality Low in Developing Countries?

## Housekeeping

- Free lecture (Oct 10)
  - Choose one: 1) firms, 2) migration, 3) policy (guest lecture)
\vfill
- Research Proposal: first draft due one week from Thursday
\vfill
- All assignments have detailed instructions and rubric
\vfill
- Start selecting replication paper

## WTP for Environment Quality in LMICs

- Crucial for policy design (Greenstone and Jack, 2013)
\vfill
- Most papers estimate $MWTP_{s}$
  - and assume market failures do not bias valuation
\vfill
- Examples of $MWTP_{s}$
  - MWTP for water filter in Ghana: \$1.80-2.40 (Berry et al. 2011)
  - MWTP for clean cookstoves in Kenya: \$12 (Berkouwer and Dean, 2022)
  - MWTP for sanitation in Cambodia: \$30 (Ben-Yishay et al., 2017)
\vfill
- Revealed Preferences: very few!
  - WTP for clean water: Kremer et al. (2011) ${\color{red} \leftarrow today}$
  - WTP for clean air: Ito and Zhang (2020)
  - \textcolor{red}{WTP for biodiversity: ???}

## Kremer et al. (2011): WTP for Clean Water in Kenya

- **Research Question:** What is household WTP for clean water in LICs?
  - What are health impacts of water quality improvements?
\vfill
- **Setting:** 1,354 households in Kenya
\vfill
- **Intervention:** Improve water quality at source for 184 natural springs
\vfill
- **Empirical Strategy:** RCT + Revealed preference valuation
\vfill
- **Results:** Clean water valued at \$USD 2.96 per hh (\$800 VSL)

## Spring Protection

- Spring water is crucial drinking water source in Africa
  - But vulnerable to contamination when people dip vessels
  - Runoff introduces animal/human waste
\vfill
- Spring protection common in Africa to improve water quality at source
\vfill
- Seal off spring at source by encasing in concrete
  - Water flows out of pipe rather than ground, avoiding contamination
\vfill
- Nothing can ``break'' $\rightarrow$ less maintenance than wells/pumps
\vfill
- How much are hh WTP for spring protection?

## Spring Protection
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "68%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/spring_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Spring Protection
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "75%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/spring_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Spring Protection
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/spring_fig3.png",
                        auto_pdf = TRUE
                        )

```

## Spring Protection
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "65%", fig.align = "center", fig.cap="Source: USAID"}

knitr::include_graphics(
                        path = "assets/spring_fig4.png",
                        auto_pdf = TRUE
                        )

```

## Study Sample
- Sample frame: list of all unprotected springs in Busia and Butere-Mumias districts
\vfill
- Treatment: 184 randomly selected for protection
  - 1,345 households selected that use springs
\vfill
- 7-8 households from each spring select from ``user list''
\vfill
- Phased design: all springs eventually protected
  - Round I (Jan-Apr 2005); 47 springs, 350 households
  - Round II (Apr-Nov 2005); 46 springs; 349 households
  - Round III (2005-2006); 91 springs; 685 households

## Data Collection

- Water quality measured at each spring based on EPA protocols
  - Measure: E.Coli (log of \# bacteria per 100ml)
\vfill
- Household survey
  - Diarrhea
  - Anthopometrics
  - Water collection behvavior
  - Socioeconomic data
\vfill
- No difference in baseline water quality between treatment/control springs

## Water Quality and Health
- Impact of spring protection on water quality:
$$W_{jt'}^{SP} = \alpha_{t} + \phi_{1}T_{jt} + X_{j}^{SP'}\phi_{2} + (T_{jt} \cdot X_{j}^{SP})'\phi_{3} + \epsilon_{jt}$$
  - where $W_{jt}$ is water quality for spring $j$ at time $t\in {0,1,2,3}$
  - $T_{jt}$ is treatment status (=1 after protection)
  - $\phi$ = impact of spring protection on water quality
\vfill
- Impact on health
$$Y_{ijt} = \alpha_{i} + \alpha_{t} + \phi_{1} T_{jt} + X_{ij}'\phi_{2} + (T_{jt} \cdot X_{ij})'\phi_{3} + u_{ij} + \epsilon_{ijt}$$

  -  where $Y_{ijt}$ is whether child $i$ had diarrhea in the past week

## Results: Impact of Spring Protection on Water Quality and Health

**Water Quality**

- Huge improvement in water quality: 66\% $\downarrow$ in E.coli
- No heterogeneity by baseline hygiene knowledge or education
  - Hypothesis: educated hh may maintain spring better
- Improvement in home water quality: 24\% $\downarrow$ in E.coli
  - Why is the effect smaller?

\vfill
**Health**

- Diarrhea incidence $\downarrow$ by 4.7pp (20\% reduction compared to control)
- No impact on child weight
- Positive effect on BMI
\vfill

**What if control households choose to use treatment springs?**

## Revealed Preference: Valuing Clean Water in Kenya

- Let valuation of water from source $j$ be $Z_{j}$
  - reflects health and non-health attributes (e.g. ease of collection)
\vfill
- $\beta_{i}$: indirect utility from protection ($T_{jt}$) at $j$ at time $t$
\vfill
- $C_{i}>0$ is time cost per minute of household $i$
\vfill
- $C_{i}D_{ij}$: Travel cost for additional trip to $j$ where $D_{i}$ is round-trip distance
\vfill
- Each trip affected by unobserved factors (weather, queue, errands, mood)

## Revealed Preference Approach
- Household indirect utility from one trip to $j$ at time $t$ is:
$$ u_{ijt} = \beta_{i}T_{jt}+Z_{j}-C_{i}D_{ij} + e_{ijt}$$
  - where $e_{ijt}$ is iid type I extreme value distributed
\vfill
- $i$ chooses $j$ over $k$ if:
$$\beta_{i}(T_{jt}-T_{kt})+(Z_{j}-Z_{k})-C_{i}(D_{ij}-D_{ik})+(e_{ijt}-e_{ikt}) \geq 0$$
- Using CDF of EV Type I error, prob. of choosing $j$ over alternatives $h$:
$$P(y_{ijt}|X) = \frac{exp(X_{ijt}^{'}\beta)}{\sum_{h}exp(X_{iht}^{'}\beta)} \equiv \rho_{ijt}$$

## Mixed Logit

- Add preference heterogeneity with mixed logit:
$$P(y_{ijt}|X) = \int_{\beta} \rho_{ijt}f(\beta)d\beta$$
  - where $f(\cdot)$ is the mixing distribution
- Maximize SLL to estimate mean and s.d. of these distributions

- **What is the object of interest?**
  <!-- - ratio of$\beta_{i}$ to coeff. on travel cost -->
  <!-- - Value of spring protection in terms of minutes spent walking -->

## WTP results
- Conditional logit estimates: how do we interpret these?
  - Coefficient on walking distance ($C_{i}D_{ij}$) = -0.055
  - Coefficient on spring protection ($\beta_{i}$) = 0.51
\vfill
- **Can $\beta_{i}$ entangle other amenity values?**

  <!-- - reduced need to bend down, lower collection times -->
  <!-- - yes, but when include perceived water quality, $\beta_{i}=0$, meaning the greater value placed on protected springs is due to impact on water quality, not other factors -->

\vfill
- Coeff. on ($T_{jt}$ $\times$ under-3 child) is positive (insignificant)
  - Households with kids value water quality more

## WTP for spring protection

- Ratio of **mixed logit** coefficients: 2.95/0.21=14 minutes

- Households value clean water by willing to walk 14 extra minutes

- Average hh takes 48 trips per week (see Descriptive Stats)
- Total extra minutes per year = 14 x 48 x 52 = 35,078 minutes
- Assuming 8-hour workday $\rightarrow$ **mean value of clean water is 73 workdays**
  - Note: in paper, it is 32.4 workdays!

## Converting WTP to dollars

**Survey evidence**

- Subset of 104 respondents
- Contingent valuation: are you WTP \$5 to walk 1 less minute? \$10?, etc
- Divide WTP (in terms of time) by money value of time
- Regress time value on household characteristics, and predict on full sample
- Mean value of time = \$USD 0.088 for 8-hr workday

\vfill

**Wages**

- \$USD 1.26 for 8-hour day (Suri, 2009)
- Time cost = 25\% of average workday wage

## Comparison with Contingent Valuation

- Households in protected spring communities asked:
  - **How much are you WTP per year to keep the spring protected?**
\vfill
- In follow up survey:
  - First: are you willing to pay 500 shillings (\$7.14)?
  - Second: are you willing to pay 1000 shillings (\$14.29)?
  - Then emphasize tradeoff (opportunity cost) and ask again
\vfill
- Almost all WTP \$7.14
  - Majority WTP \$14.29 even after learning about tradeoff
  - Starting value matters!

<!-- ## Distribution of Household Values of Clean Water (Mixed Logit) -->
<!-- ```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"} -->

<!-- knitr::include_graphics( -->
<!--                         path = "assets/kremer2011_fig1.png", -->
<!--                         auto_pdf = TRUE -->
<!--                         ) -->

<!-- ``` -->
<!-- - Why is stated preference values larger? Why is CV distribution wider? -->

## Health Valuation and VSL

- Spring protection averts **3.2 diarrhea cases per household-year**
  - (0.047 cases/child-week) * (1.3 children under 3 per hh) * (52 weeks/year)
\vfill
- Annual WTP for spring protection = 32.4 workdays/year
  - Assume this value captures benefits from averted diarrhea
\vfill
- If protection prevents 3.2 cases/yr, valued at 32.4 workdays/yr, then
\vfill
$\text{WTP/case averted} = \frac{\text{Total Workdays}}{\text{Diarrhea Cases Averted}} = \frac{\text{32.4 workdays}}{\text{3.2 cases}} = \text{10.1 workdays/case}$

## Health Valuation and VSL

- Assume spring protection $\downarrow$ diarrhea morbidity \& mortality proportionally

- Using fact that 862 diarrhea cases $\rightarrow$ 1 death (literature):

- WTP to prevent 862 diarrhea cases (1 death) i.e. VSL:
$$ \text{Total WTP/death averted}= \text{10.1 workdays} \times \text{862 cases} \approx \text{8,700 workdays}$$

**VSL = \$0.088 $\times$ 8700 $\approx$ \$USD 770**

**In USA 2013: VSL $\approx$ \$USD 8.5 million (EPA)**

## Discussion

- What do you think about approach?
- What do you think about time valuation?

## 
\centering
\Huge Lecture 8

\large Why is WTP for Environmental Quality Low in LMICs?

## Housekeeping

- First draft due one week from today

- Select your replication paper

<!-- ## Today's lecture -->

<!-- - Why is willingness to pay so low in developing countries? -->

<!-- - One reason: **Revealed $MWTP_{e} \neq MWTP_{e}$ from first best** -->

<!-- - Market failures create a wedge -->

<!-- - Today: experiments to alleviate frictions -->

## Energy efficiency uptake in LMICs
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/bd2022_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Market Failures in Development Economics

- Market failures can shape observed $MWTP_{e}$
  - And, therefore, environmental policy

\vfill
- Main frictions: information, credit, risk, or property right imperfections
  - Frictions warp $MWTP_{e}$, resulting in lower EQ

\vfill
- **Revealed preference** $MWTP_{e}$ assumes agent knows payoffs
  - Misinformation common in developing countries
  - Due to literacy, education, information access, state capacity, etc.

## Credit Market Failures

- Stem from difficulty writing/enforcing contracts in developing countries
\vfill
- Costs of monitoring borrowers + enforcing repayment leads to high interest rates and credit rationing (Conning and Udry, 2007)
\vfill
- Agents cannot pay upfront for clean technology that has future payoffs
  - Forego future improvements in EQ, income, and health
\vfill
- **Determinants of liquidity therefore confound measured $MWTP_{e}$**

## Missing Risk Markets

- Lower $MWTP_{e}$ if payoffs are uncertain and insurance unavailable
\vfill
- Worse EQ can increase var(Y)
\vfill
- Missing risk markets $\uparrow$ exposure to risk and $\downarrow$ measured $MWTP_{e}$
  - Liquidity: prioritize income for emergencies, instead of clean water
  - Discounting: undervalue investment in EQ because uncertain about whether will be able to enjoy benefits (discount future more heavily)

## Evidence on the effect of market failures on MWTP

- What is the ideal experiment?

- How do you distinguish between measured and actual $MWTP_{e}$?

- Would random variation help?

##

\begin{quote}
`` there is hardly a more important topic
for future study than developing revealed
preference measures of $MWTP_{e}$ that capture
the aesthetic, health, and/or income gains
from environmental quality''
\vspace{0.5cm}\\
- Envirodevonomics (Greenstone and Jack, 2013)
\end{quote}

## Berkouwer and Dean (2022): Frictions in WTP for Clean Air

- **Research Question:** What are the barriers to adopting clean cookstoves?

- **Setting:** Nairobi, Kenya
  - Test whether credit constraints and inattention to cost-savings lead to underadoption

- **Data:** 1000 households
  - Measure WTP with BDM mechanism

- **Research Design:** RCT with 1000 households to quantify WTP drivers
  - Three-by-three experimental design
  - Cross-randomize two credit and two attention treatments

- **Results:** 1) Stove generates \$237 in charcoal savings, 2) yet only \$12 WTP, 3) credit constraints are the main friction preventing takeup

## Key Idea

WTP for enviromental quality $\rightarrow$ WTP for improved stove

## Experimental Setting
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/bd2022_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Model of Adoption
- Consider agent evaluating an energy efficient technology (costs $P_{E}$)
- Adopt iff sum of savings ${\color{blue} S_{t}}$ exceed costs today:
$$\underbrace{u(c_0) - u(c_0 - P_E)}_{\text{cost of adoption today}} < \underbrace{\sum_{t=1}^{T} \delta^t [u(c_t + {\color{blue} S_t}) - u(c_t)]}_{\text{future benefits}}$$
- Agent can take loan ${\color{red} L}$ and pay back in payments ${\color{red} r_{t}}$:
$$\underbrace{u(c_0) - u(c_0 - P_E+{\color{red} L})}_{\text{cost of adoption today}} < \underbrace{\sum_{t=1}^{T} \delta^t [u(c_t + {\color{blue} S_t} - {\color{red} r_{t}}) - u(c_t)]}_{\text{future benefits}}$$
- WTP $P^{*}$ is when agent is indifferent
$$u(c_0) - u(c_0 - P^{*}+{\color{red} L}) = \sum_{t=1}^{T} \delta^t [u(c_t + {\color{blue} S_t} - {\color{red} r_{t}}) - u(c_t)]$$

## Model of Adoption

**Market Failure Tests**

- Agent may be **inattentive** to energy savings by ${\color{blue} \theta_{s}} \in [0,1]$
  - concentration bias: less attentive to small payments ${\color{red} r_{t}}$
  - Time inconsistency: inattentive to future loan payments ${\color{red} \theta_{c}} \in [0,1]$
\vfill
- Agent may be **credit constrained** ${\color{red} L \leq \bar{C}}$

## Empirical Tests

Main outcome variable: WTP $P^{*}$
$$u(c_0) - u(c_0 - P^* + {\color{red} L}) = \sum_{t=1}^{T} \delta^t \left[ u(c_t + {\color{blue} \theta_s S_t} - {\color{red} \theta_c r_t}) - u(c_t) \right] \quad \text{s.t.} \quad {\color{red} L \leq \overline{C}}$$

1. Randomize attention to energy savings $\theta_s$ \hfill $\frac{\partial P^*}{\partial \theta_s} > 0$

2. Randomize access to credit $\overline{C}$ \hfill $\frac{\partial P^*}{\partial \overline{C}} > 0$
    - Randomize into \( N \) payments of size \( r_t \) \hfill $\frac{\partial^2 P^*}{\partial \overline{C} \partial r_t} < 0$

    - Randomize attention to costs \( \theta_c \) \hfill $\frac{\partial^2 P^*}{\partial \overline{C} \partial \theta_c} < 0$

    - Measure time-inconsistency (TI) \hfill $\frac{\partial^2 P^*}{\partial \overline{C} \partial TI} > 0$

## Research Design: Cross-Randomize Credit and Attention
- Each cell contains number of respondents

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/bd2022_fig3.png",
                        auto_pdf = TRUE
                        )

```

## Credit: Control and Treatment Groups

- **Control:** pay (subsidized) price today, receive stove today

- **Treatment:** receive stove today, pay w/n 3 months
  1. Pay by 12 weekly deadlines
  2. Pay by 3 monthly deadlines

- SMS reminders if miss deadlines

- Interest rate of 1.16\% per month, 14.8\% per year

## Attention: Control and Treatment

- **Control:** Informed that stove reduces charcoal use by 50\%
  - Informed of \$ equivalent of savings
\vfill
- **Treatment 1:** Attention to energy savings
  - Respondent writes how much they think they'll save
  - Writes down how they will use the savings
\vfill
- **Treatment 2:** Attention to energy savings minus costs

## Attention to Savings

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/bd2022_fig4.png",
                        auto_pdf = TRUE
                        )

```

## Attention to Savings minus Costs

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/bd2022_fig5.png",
                        auto_pdf = TRUE
                        )

```

## Elicit WTP in Each Group

- **Becker-DeGroot-Marschak (BDM) Mechanism**
  - Field worker puts random price $P_{i} \in (0.01-29.99)$ in envelope
  - Elicit max $WTP_{i}$ from respondent using binary search
  - Open envelope
  - WTP is binding: pay $P_{i}$ iff $P_{i} \leq WTP_{i}$
\vfill
- Benefits
  - Generates **incentive-compatible** behavior! Why?
  - Adoption is random, conditional on WTP

\vfill
**BDM mechanism elecits WTP AND generates random stove adoption**

## Stove reduces charcoal spending by \$2.28 per week!
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/bd2022_fig6.png",
                        auto_pdf = TRUE
                        )

```

- Large effects: \$120 savings per year
- Given \$40 cost of stove, this is **300\% rate of return!**

## Yet, demand is low

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/bd2022_fig7.png",
                        auto_pdf = TRUE
                        )

```

## No evidence of inattention to energy savings!

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/bd2022_fig8.png",
                        auto_pdf = TRUE
                        )

```
- Information frictions may not be attenuating WTP

## Access to credit doubles WTP!
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/bd2022_fig9.png",
                        auto_pdf = TRUE
                        )

```
- credit increases WTP by \$13 (104\% relative to control)
<!-- - credit push demand curve to breakeven point (i.e. use savings to pay for stove) -->

## Other Reasons why WTP is low

- Information access: Baylis et al. (2024), Jalan and Somanathan (2008), Bennear et al. (2013)
\vfill
- Credit: Ben-Yishay et al. (2017)
\vfill
- Property rights: Ali et al. (2014), Abman and Carney (2020), Bellemare et al. (2018), Noack and Costello (2024)

## Discussion

- Market frictions affects many things $\rightarrow$ endogeneity

- Partial vs. general equilibrium

- Cannot separate measurement from actual distortions
  - Compare across methods more/less affected by frictions?