### 2024-11-11 lectures/module5/module5.Rmd

---
title: "| Environmental and Development Economics\n| Module 5 - Environmental Policy Design
  \n"
author: |
  | Raahil Madhok
  | UMN Applied Economics
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Montpellier
  slidy_presentation: default
  ioslides_presentation: default
classoption: aspectratio=169
header-includes:
- "\\AtBeginDocument{\\title[Module 5 - Environmental Policy Design]{Environmental and Development
  Economics \\\\ Module 5 - Environmental Policy Design in LMICs}}"
- \input{header.tex}
- \usepackage{gensymb}
---

## 
\centering
\Huge Lecture 8 (continued)

\large Environmental Policy Design in LMICs: Monitoring and Enforcement

## Housekeeping

## Rich countries have resolved many environmental problems
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/regulation_fig1a.png",
                        auto_pdf = TRUE
                        )

```

## Within-country: Richer countries effectively use environmental regulation
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "50%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/regulation_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Role of the state in improving environmental quality

- With perfectly functioning markets, agent sets:

$$MWTP_{e} = \frac{\partial c_{e}}{\partial \Delta e}$$

- This yields **first-best outcome**
\vfill
- For policy design, social planner sets optimal $e$
  - **Assumption 1:** Assume they know true $MWTP_{e}$ (last week)
  - **Assumption 2:** Costless technology to raise revenue
\vfill
- **Assumption 2** fails in LMICs $\longleftrightarrow$ $\frac{\partial c_{e}}{\partial \Delta e}$ too high!

## Optimal Environmental Policy Design 101

:::::::::::::: {.columns}
::: {.column width="45%"}

- LMICs on left of equilibrium
\vspace{0.3cm}
- Optimal to abate until MAC=MB
\vspace{0.3cm}
- Should be easy since MAC low
\vspace{0.5cm}

**puzzle:  $\frac{\partial c_{e}}{\partial \Delta e}$ is high!**

- why?

:::
::: {.column width="45%"}
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/mac.png",
                        auto_pdf = TRUE
                        )

```
:::
::::::::::::::

## Today

- **Guiding question:** How do we design optimal environmental policy in LMICs?

\vfill
- Today's focus: challenges for designing low MAC policies
  - Perverse incentives (Duflo et al., 2013)
  - Spillovers and unintended consequences (Englander 2023)

\vfill
- Next time: other barriers to efficient policy design
  - Trust
  - Knowledge

## Background: Poor countries DO have environmental policies

```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/regulation_fig3.png",
                        auto_pdf = TRUE
                        )

```

## Background: India

- Command-and-control regulation is the norm
\vfill
- "Command" is easy, "control" is costly
\vfill
- State Pollution Control Board (SPCB) enforces regulatons
  - State capacity is weak
  - Enforcement is very costly $\rightarrow$ less monitoring
\vfill
- **Role of third-party auditors**

## Duflo et al. (2013): Truth-Telling by Third-party Auditors

- **RQuestion:** How much do firms skirt environmental policy in developing countries?
  - Can realigning incentives of auditors $\uparrow$ compliance with environmental norms?

- **Setting:** Pollution audit system in Gujarat, India

- **Data:** 473 firms; audits, backchecks, and pollution monitoring

- **Research Design:** RCT
  - Treatment: random auditor assignment with fixed payment

- **Results:** 1) widespread corruption in status quo, 2) more truth-telling in treatment, 3) firms $\downarrow$ pollution

<!-- ## Background -->

<!-- - Third party auditing used to monitor firm regulatory compliance -->
<!--   - Financial sector: credit ratings -->
<!--   - Auto: compliance with vehicle standards -->
<!--   - Food retail: health inspector -->
<!--   - **Environmental:** compliance with national/state laws -->

<!-- - Feature: auditor chosen by, paid by, and reports to audited firm -->

<!-- - **Incentive to shade/falsify reports!** -->
<!--   - Corrupt information provision and undermine regulation -->

## Context: Gujarat, India

- One of India's fastest growing states
  - Home to the two post polluted industrial clusters

- Pollution standards set nationally, enforcement is state-level
  - GPCB responsible for enforcing Air Act

- GPCB uses regulates 20,000 plants

- Main regulatory instrument: plant-level inspections + third-party audits

## Environmental Audits in Gujarat

- Polluting firms submit yearly environmental audit

- \textcolor{red}{Audit conducted by audit firm hired and paid by the plant}

- Auditors visit plant three times/year to measure pollution
  - Submit audit report to GPCB

- (On paper) checks and balances on auditors
  - Max 15 plants per year
  - Can audit a plant at most 3 years in a row
  - Auditors with inaccurate reports can be decertified

## Environmental Audits in Gujarat

- (On paper) checks and balances on plants
  - Failure to submit report punishable by closure/disconnection
  - Noncompliance audit report punishable by closure/disconnection
\vfill  
- Despite safeguards, system functions poorly
  - GPCB believes inaccurate reporting renders audits useless for enforcement
\vfill
- Evidence of auditor shopping: Plants pay $\approx$ INR 24,000 (\$300)
  - Significantly less than true costs of conducting audit

## Experimental Design: N=473 plants

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig5.png",
                        auto_pdf = TRUE
                        )

```

## 
\centering
\Huge Lecture 9

\large Environmental Policy Design in LMICs

## Housekeeping

- First draft due on Thursday
  - Peer review due one week later (2 pages)
\vfill
- Proposal presentation (Oct 15th and 17th)
  - Consider drafting it as you write your first draft
  - Rubric added in github
\vfill
- Replication (3 pages, due Oct 21st)
\vfill
- Minor syllabus updates
  - State capacity in-class presentation
\vfill
- Migration lecture readings will be posted this week

## Model: Set Up

- **Three Players**: plant, auditor, and  regulator.
\vfill
- Plants face the decision to:
  - **Abate pollution**: Incurs a cost, but avoids penalties.
  - **Falsify reports**: Pay auditor to underreport pollution and avoid abatement.
\vfill
- Auditors considers:
  - Payment from plant
  - Risk of disbarment by regulator if found cheating.
\vfill
- Regulator: reduce pollution by enforcing penalties and disbarring false auditors

## Plant's Decision and Cost Function

- The plant's goal is to minimize total costs:

$$S - 1\{Abate\} \cdot c - w - 1\{\text{Fine Imposed}\} \cdot f $$

- Where:
  - S = Value of production
  - c = Cost of abating pollution
  - w = Payment to auditors for the report
  - f = Fine imposed if non-compliance is detected

- **Abatement**: Reduces emissions from $p_{h}$ to $p_{l}$, at cost $c \sim G(\cdot)$

- **Non-compliance**: Plant chooses $w(\hat{p}|p)$ to minimize costs
  - where $\hat{p}$ is auditor report, and $p$ is true pollution

## Auditor’s Payoff and Reporting Incentives

- Auditor's payoff function is:

$$w - a + E[V(\hat{p}|p)]$$

- Where:
  - $w$ is payment for performing audit
  - a = Cost of performing the audit
  - $E[V(\hat{p}|p)]$ = continuation value of future business
  - Assume leniency increases EV: $V_{lh} = V(p_{l}|p_{h})>V_{hh} =V_{0}$
\vfill
- Auditors may have incentive to falsify reports to maintain business
  - Higher \( V_{lh} \), more attractive for auditors to underreport
  - $E[V(\hat{p}|p)]$ incorporates likelihood of disbarrment

## Regulator Problem

- Reviews reports $\hat{p}$ and may impose $f$ on plant or disbar auditor

- Let $q$ = probability regulator observes true $p$ (backchecks)

- Decide whether to impose $f$

- Continuation value of disbarred auditor is zero

- Regulator only goal: minimize pollution

## Equilibrium: Proposition 1

Unique subgame-perfect equilibrium with regulation:

-  **Plants Abate**: If \( c \leq \min\{ f, qV_{lh} - \Delta V \} \).
-  **Underreporting**: If \( qV_{lh} - \Delta V < \min\{ c, f \} \).
-  **Defy Regulation**: If \( f < \min\{ c, qV_{lh} - \Delta V \} \).

\vfill

**Intuition**

- **Abatement**: when cheaper than paying fines or paying auditors to cheat

- **Underreport**: when paying auditors to cheat cheaper than abatement and fines

- **Defiance** when both abatement and bribing more expensive than risking fines

## Equilibrium: Proposition 2 - Backchecks

**Status Quo (control group)**

- $q \rightarrow 0$ in status quo
- Negative cost of false reports: $qV_{lh} - \Delta V$
- Result: Widespread underreporting and no incentive for abatement

\vfill

**With Backchecks and Incentive Pay (Proposition 2)**

- Backchecks: $\uparrow$ q s.t. $qV_{lh} - \Delta V > f$ induces fraction $G(f)$ of plants to abate
- Incentive pay: Makes underreporting more costly than fine if $q(V_{lh}+B)-\Delta V > f$

\vfill

**Intuition**:

- Backchecks increase $q$, raising expected cost of false reporting
- When $qV_{lh} - \Delta V > f$, some plants prefer abatement over risking detection.
<!-- - Higher $q$ makes underreporting less attractive, shifting plants towards compliance -->

## Random Assignment of Auditors

- **Flat fee:** Payment unconditional on report outcomes: $w({\hat{p}|p})=w(p)$
  - Plant cannot bribe auditor, nor can they refuse to pay

- Auditors only consider reputation $\rightarrow$ increase in $q$ for plants to abate is smaller

- Reduces continuation value of notoreity, $\Delta V$

\vfill

**Proposition 3**: Rise in $q$ s.t. $qV_{lh}\geq \Delta V$ induces share $G(f)$ to abate

\[
\text{If } qV_{lh} \geq \Delta V \implies \text{Plants will abate if } c \leq f
\]
\vfill

- **Intuition**: If treatments are strong enough s.t. $q(V_{lh}+B) > \Delta V$, then $G(f)$ plants will abate and rest will report higher $p$ and risk fine.

<!-- - Incentive Pay increases \( V_{lh} \) for accurate reporting s.t. false reporting less attractive -->

## Model Summary

- **Higher \( q \)** through backchecks increases risk of detection for false reporting

- **Random assignment** and **centralized payment** $\downarrow$ auditors-plant dependency

- **Incentive Pay** boosts auditors' continuation value for truthful reporting

- Treatment makes false reporting costlier for auditors and plants

- Plants either abate pollution or risk fines, depending on abatement costs

- The equilibrium shifts towards increased compliance and truthful reporting

## Data

- **Audit Reports (2009 & 2010)**: water and air pollution
   - Collected from auditors in both treatment and control groups
\vfill
- **Backchecks**: independent *unannounced* verification of pollution levels
   - Conducted randomly on 20% of treatment plants in 2009 and 2010
   - Midline backchecks in 2010 for both treatment and control groups.
\vfill
- **Endline Survey (2011)**:
   - Measured actual pollution outcomes after the audit intervention
\vfill
- **Administrative Data from GPCB**:
   - Regulatory actions, violations, and enforcement measures
   - Fines, closures, and disconnections imposed on plants

## Results: Status Quo Pollution
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig3.png",
                        auto_pdf = TRUE
                        )

```

## In Regression Form
$$ 1\{Compliant\}_{ij} = \beta_{1}1\{AuditReport\} + \alpha_{r} + \epsilon_{ij}$$
\pause
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Treatment Effects: Truth Telling About Plant Compliance

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig1.png",
                        auto_pdf = TRUE
                        )

```

## In Regression Form
\vspace{-0.5cm}
$$ 1\{Compliant\}_{ij} = \beta_{1}1\{AuditReport\} \times T_{j} + \beta_{2}1\{AuditReport\} + \beta_{3}T_{j} + \alpha_{r} + \epsilon_{ij}$$
\pause
\vspace{-0.5cm}
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "65%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig4.png",
                        auto_pdf = TRUE
                        )

```
- Treatment increases truth-telling: 19pp less cheating in treatment

## Which firms are getting told on?

- Convert pollution readings to deviation from regulatory standard
\vfill
- Readings grouped into 40 different 0.05$\sigma$ width bins
\vfill
- 40 regressions: regress dummy for each bin on treatment indicator
  - Negative coefficient: treatment auditors less likely to report values in that bin
\vfill 
- Result: treatment reduces the mass just beneath the standard
  - **More likely** to report values above standard

## Density of audit report pollution distribution
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "55%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig6.png",
                        auto_pdf = TRUE
                        )

```
- Readings by treatment auditors 20\% less likely to be in bin just below standard compared to control readings

## Do plants reduce emissions?
- From model: treatment induces $G(f)$ plants abate by $\Delta$ incentives
- Regress plant pollution on treatment + region FE in endline (cross section)
\pause
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig8.png",
                        auto_pdf = TRUE
                        )

```
- Treatment plants $\downarrow$ pollution by 0.21$\sigma$

## Which plants are reducing pollution?
- Quantile regression: $Q_{y_{ij}|X_{j}}(\tau) = \beta T_{j} + \alpha_{r} + \epsilon_{ij}$
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "45%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig9.png",
                        auto_pdf = TRUE
                        )

```
- Clean-up driven by dirty plants (>80th pctile)
- Driven by monotonic increase in severity of GPCB fines w/ non-compliance

## Policy 
- Eliminating conflict of interest led to policy change
- Scaled up to state level since 2015

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig10.png",
                        auto_pdf = TRUE
                        )

```

## Thoughts?

##

\Large Other challenges for designing low MAC policies?

## Barriers to Optimal Policy Design

- Recall, goal of social planner is to internalize externalities
\vfill
- Objective is to set $MWTP_{e} = \frac{\partial c_{e}}{\partial \Delta e}$ using **regulation**
\vfill
- We saw that low state capacity creates a wedge by keeping $\frac{\partial c_{e}}{\partial \Delta e}$ high
  - This is true even if market failures absent
\vfill
- **Top down:** $\uparrow$ state capacity **directly** has large benefits (more next week)
\vfill
- **Bottom up:** what if we instead change citizens' trust in the state?

## Jack et al. (2013): Money not to Burn

- **Research Question:** Can PES contracts curb crop burning?
\vfill
- **Setting:** North India; weak institutions for contract efficacy
\vfill
- **Data:** 171 villages in North India
\vfill
- **Design:** RCT with two treatments: 1) standard PES 2) partial payment upfront
\vfill
- **Results:** Upfront PES has 10pp higher compliance than standard PES
  - Takeaway: trust is an important barrier for efficient environmental policy design

## Textbook externality problem
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/jack2023_fig5.png",
                        auto_pdf = TRUE
                        )

```
## Background

- Crop  burning major source of pollution in Punjab
\vfill
- Burning rice stalks post-harvest $\rightarrow$ 40% of winter pollution in Delhi
  - Severe health risks
\vfill
- Ban on burning since 2015, with fines for violations, but weak enforcement
  - Fine: Rs. 2,500 to Rs. 15,000 based on landholding
\vfill
- Subsidies for residue management equipment exist
  - insufficient to cover farmers' costs (still cheaper to burn)

## PES Contracts

- PES contracts pay farmers for environmentally friendly practices
\vfill
- Goal: increase the private costs of burning to incentivize ``good'' behavior
\vfill
- Advantage: targets outcome not inputs
  - i.e. agnostic about technology, as long as abatement cost $<$ payment
\vfill
- Challenge: trust and liquidity constrain standard PES contracts in LMICs
  - Treatment: pay 25\% or 50\% upfront

## RCT Design
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/jack2023_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Data

- Baseline survey: 1,668 farmers (N=171 villages) from two districts in Punjab
\vfill
- Data: Demographics, income, trust in institutions, and CRM equipent
  - Farm plot shapefiles for linking with satellite imagery
\vfill
- Monitoring: field visits, random spot checks, satellite imagery

## Estimation

$$ y_{ij} = \alpha + \beta StandardPES_{j} + \gamma UpfrontPES_{j} + \psi X_{j} + \epsilon_{ij}$$

- $y_{ij}$ = outcome for farmer $i$ in village $j$
- $X_{ij}$ are strata fixed effects
- Subtreatments are pooled
- $\beta$: effect of standard PES relative to control
- $\gamma$: effect of upfront PES relative to control

## Results: Contract Compliance
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/jack2023_fig2.png",
                        auto_pdf = TRUE
                        )

```
- Standard PES: 8.5\% complied with contract
- Upfront PES: 18\% complied with control --- trust and liquidity important!

## Other Results

- Relative to control, upfront PES $\uparrow$ non-burning by 7pp (85\%)
\vfill
- Most farmers still burned...
  - Cost of alternatives to burning still higher than PES payment
\vfill
- Standard PES insignificant in column 2 (inframarginal)
  - farmers who complied would not have burned anyway

## Mechanisms: Why do upfront PES increase compliance?

- Hypothesis: 1) distrust in conditional payment, 2) limited cash to rent CRM

```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/jack2023_fig3.png",
                        auto_pdf = TRUE
                        )

```
- Upfront PES no better for farmers w/ liquidity constraints or distrust

## Mechanisms: Why do upfront PES increase compliance?

```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/jack2023_fig4.png",
                        auto_pdf = TRUE
                        )

```
- Upfront farmers 7pp more likely to trust that payment will be made
- CRM cash shortages not important for either upfront or standard farmers
<!-- - Upfront amount may have been too low to ease constraint -->

## Discussion

- Limited short-term capital and distrust in institutions can limit effectiveness of standard PES contracts

- May generalize to conditional CCTs more broadly

- Thoughts?

##
\huge Other barriers to efficient policy design

\large Unintended Consequences: Spillovers

## Englander (2023): Information \& Spillovers in Conservation Policy

- **Research Question:** What are the direct and spillover effects of targeted conservation policy?
\vfill
- **Setting:** Temporary closures in Peru's anchoveta fishery
\vfill
- **Data:** location, time, and \# of fish caught per boat
\vfill
- **Research Design:** ``Potential'' closure algorithm
\vfill
- **Results:** direct effect: lower catch in closed areas
  - indirect effect: big positive spatial spillovers
  - net effect: policy backfires and $\uparrow$ catch!

## Targeted policies are common forms of incomplete regulation

- Reduce externality by targeting highest MD places
  - Road congestion pricing
\vfill
- Incomplete regulation $\rightarrow$ spillovers (impacts on non-targeted units)
\vfill
- Direction and magnitude of spillovers varies

## Peru's Anchoveta Fishery

:::::::::::::: {.columns}
::: {.column width="45%"}

\vspace{1cm}
- World's largest (tons caught)
\vspace{0.3cm}
- 8\% of global marine catch
\vspace{0.3cm}
- \$2 billion export revenue per year
\vspace{0.5cm}
- 730 active vessels per season

:::
::: {.column width="50%"}
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig1.png",
                        auto_pdf = TRUE
                        )

```
:::
::::::::::::::

## Catching juveniles is higher marginal damage than adults

```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "40%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Spatial Closures

- Regulator temporarily closes area of ocean when \% of juveniles $>$ 10\%

- Regulator exercises discretion; sometimes no closure

- Vessels report catch with electronic logbooks
  - Time, tons, and \% juvenile from each set
  - **Challenge:** fishermen underreport \% juvenile
  - **Solution:** third-party inspection data

- **Closure rule:** observe \% juvenile
  - Draw rectangle around overfished sets
  - Close area inside box
  - Announce on website, email, whatsapp

## Reported Size Distribution (example set)
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig3.png",
                        auto_pdf = TRUE
                        )

```

## Corrected Size Distribution (example set)
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig4.png",
                        auto_pdf = TRUE
                        )

```

## Counterfactual Closures

- Mimic the closure rule (potential closures)

- Intersect potential closures (N=973) w/ actual declared closures (N=410)

- **Control for length distribution from which \% juveniles drawn**
  - Comparable distribution of fish sizes

- Similar potential closures, regulator believes different

## Treatment and Control Groups
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig5.png",
                        auto_pdf = TRUE
                        )

```

## Identifying variation

- Comes from controlling for length distribution
\vfill
- Compare potential closures that by chance had higher \% juvenile draws (and were thus closed by regulator) to potential closures that by chance had lower \% juvenile draws (and were thus not closed by regulator)
\vfill
- Key to identification: fishers report \% juvenile catch from 200 anchoveta out of several million in a set

## Empirical Strategy: Treatment Window for Spillovers
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig9.png",
                        auto_pdf = TRUE
                        )

```

## Empirical Strategy

$$ JuvenileCatch_{ist} = \alpha_{st} + \beta_{st}TreatFraction_{ist} + \sum_{l} \eta_{l}Prop_{il} + \Gamma X_{i} + \sigma_{wg} + \delta_{d} + \epsilon_{ist}$$

- i = potential closure
- s = spatial unit
- l = half-cm length interval
- w = two weeks 
- g = grid cell
- d = day

## Event Study
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig6.png",
                        auto_pdf = TRUE
                        )

```

## Main Results
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig7.png",
                        auto_pdf = TRUE
                        )

```

## Mechanism: Information Signal

- Game theoretic model
- Vessel chooses where to fish to max. expected profits
- Policy increases catch if:
  - Closures are +'ve signal relative to congestion costs
  - \% juvenile sufficiently high near closures
\vfill

- **Predictions:** 
  - Ex-post value of information competed away in equilibrium
  - Expected profit would be higher near closures if not for signal
  - Vessels receiving bigger info shocks have larger treatment effects

## Results: Value of positive signal is competed away

$$Tons_{vjk} = \beta_{1}Near_{vjk} + \beta_{2}DistToShore_{vjk} + \delta_{vj} + \gamma_{d} + \alpha_{jg} + \epsilon_{vjk}$$

where v=vessel, j=season, k=set, d=day, g=gridcell
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig8.png",
                        auto_pdf = TRUE
                        )

```

## Thoughts?

- Targeted policy backfires due to positive spillovers
- Policy implicitly signals info about non-targeted units